%!TEX root = ../rules-working.tex
%LTeX: enabled=false

\begin{onecolumntablefloat}
\begin{onecolumntable}

\tablecaption{table:acceleration}{Accel/Decel Point Summary.}

\begin{tabularx}{\linewidth}{C}
\toprule
Accel Point Summary\\
\midrule
\begin{itemize}
    \item Aircraft power = + Variable.
    \itemdeletedin{2A}{2A-steep-dives/2A-unloaded-dives}{Steep or Unloaded dive = +0.5 per level initially, then +1.0 per level.}
    \itemdeletedin{2A}{2A-steep-dives/2A-unloaded-dives}{Vert. dive = +1.0 per lvl.}
    \itemaddedin{2A}{2A-steep-dives/2A-unloaded-dives}{Steep dive, vertical dive, or unloaded = 1.0 AP per level.}
\end{itemize}
\\
\midrule
Decel Point Summary\\
\midrule
\begin{itemize}
    \item Turning = Variable.
    \item Sust. climb = 0.5 per lvl.
    \item \changedin{2A}{2A-zoom-climbs}{Zooms = 1.0 per level initially, then 1.5 per lvl.}{Zoom climb = 1.0 DPs per level.}
    \item Vert. climb = \changedin{2A}{2A-vertical-climbs}{2.0}{1.5} per lvl.
    \item \changedin{2A}{2A-spbr}{Speed brake usage = 1.0 per 0.5 speed lost.}{Speedbrakes = up to the maximum value on the ADC.}
    \itemaddedin{2A}{2A-idle}{Idle power = the DP value on the ADC.}
    \item 1.0 if Idle or Normal Pwr.\ and above cruise speed\addedin{2A}{2A-cruise}{\ for the configuration}.
    \itemdeletedin{2A}{2A-sustained}{\changedin{1C}{1C-apj-23-errata}{Sustained turns and rolls 1.0 each, or 2.0 if HBR.}{Sustained turns 1.0 each or 2.0 if HBR.}}
    \itemaddedin{2A}{2A-sustained}{Sustained turns = 1.0 DP per \degrees{30} facing change in the second and subsequent facing changes (0.5 if LBR and 1.5 DP if HBR).}
    \itemaddedin{1C}{1C-apj-23-errata}{Sustained rolls 1.0 each}.
\end{itemize}
\\
\bottomrule
\end{tabularx}

\end{onecolumntable}
\end{onecolumntablefloat}
