%!TEX root = ../rules-working.tex
%LTeX: enabled=false

\onecolumn
\begin{landscape}

\begin{fullwidthtable}
\tablecaption{table:helicopter-data}{Helicopter Data.}
    

\footnotesize
\setlength{\tabcolsep}{0.7em}
\begin{tabularx}{\linewidth}{Lccccclllrcccccccccccc}
\toprule
&
\multicolumn{5}{c}{Aircraft Data}&
&
\multicolumn{5}{c}{Gun}&
&
\multicolumn{4}{c}{ECM}&
\multicolumn{5}{c}{Technology}
\\
\cmidrule(lr){2-6}
\cmidrule(lr){8-12}
\cmidrule(lr){14-17}
\cmidrule(lr){18-22}
Type&
\vertical{HSA}&
\vertical{Visibility}&
\vertical{Size}&
\vertical{Vulnerability}&
\vertical{Blind Arc}&
\multicolumn{1}{c}{\vertical{Ground Unit Data}}&
\multicolumn{1}{c}{\vertical{Type}}&
\multicolumn{1}{c}{\vertical{Arc}}&
\multicolumn{1}{c}{\vertical{Ammunition}}&
\vertical{Hit}&
\vertical{AtA/AtG Damage}&
\vertical{Bomb System}&
\vertical{DDS}&
\vertical{IRM Suppression}&
\vertical{RWR}&
\vertical{DJM}&
\vertical{TV/IR Optics}&
\vertical{Laser Spot Tracker}&
\vertical{Laser Designator}&
\vertical{HMS}&
\vertical{MMS}
\\
\midrule

\multicolumn{22}{c}{US Helicopters}\\
\midrule

\tablerowaddedin{2A}{2A-helicopter-hh-53c}{HH-53C&
N&6&\plus{0}&\plus{2}&---&4-12&
\binarymultiply{7.62 mm GAU-2}{3}&LSD/RSD/RD&8&1/1/1&\phantom{**}2/2**\phantom{}&
---&---&\plus{0}&---&---&---&---&---&---&---
}

AH-1J Cobra&
Y&4&\plus{1}&\plus{1}&\arcminus{30}&5-8&
20 mm M197&\arcplus{150}&6&5/3/2&\phantom{**}3/4*\phantom{*}&
Ballistic&\binarymultiply{A}{1}&\plus{2}&A&---&---&---&---&---&---
\\

AH-64 Apache&
Y&5&\plus{0}&\plus{2}&\arcminus{30}&5-8&
30 mm M230&\arcplus{150}&6&5/3/2&\phantom{**}6/6\phantom{**}&
Computed&\binarymultiply{A}{2}&\plus{3}&A&B&Y&Y&B&Y&---
\\

500MD&
---&4&\plus{1}&\minus{1}&\arcminus{30}&3-6&
---&---&---&---&---&
Ballistic&\binarymultiply{A}{1}&\plus{1}&---&---&---&---&---&---&---
\\

OH-58D&
---&4&\plus{1}&\minus{2}&\arcminus{30}&3-6&
---&---&---&---&---&
Ballistic&\binarymultiply{A}{1}&\plus{2}&A&---&Y&Y&B&---&Y
\\

\midrule
\multicolumn{22}{c}{Soviet/Russian Helicopters}\\
\midrule

Mi-24 (Hind-A)&
---&6&\plus{0}&\plus{2}&\arcminus{60}&6-12&
12.7 mm A-12.7&\arcplus{150}&4&2/1/--&\phantom{**}1/2**&
Ballistic&---&\plus{2}&---&---&---&---&---&---&---
\\

Mi-24D (Hind-D)&
---&6&\plus{0}&\plus{2}&\arcminus{60}&6-12&
12.7 mm Yak-B&\arcplus{150}&4&3/2/--&\phantom{**}2/4**&
Ballistic&\binarymultiply{A}{2}&\plus{2}&---&---&---&---&---&---&---
\\

Mi-24P (Hind-F)&
---&6&\plus{0}&\plus{2}&\arcminus{60}&6-12&
23 mm GSh-23&Fixed&4&3/2/1&\phantom{**}4/4\phantom{**}&
Ballistic&\binarymultiply{A}{2}&\plus{2}&B&---&---&---&---&---&---
\\

Mi-28 (Havoc)&
Y&6&\plus{0}&\plus{3}&\arcminus{30}&6-12&
30 mm 2A42&\arcplus{150}&5&3/2/1&\phantom{**}5/6\phantom{**}&
Computed&\binarymultiply{A}{2}&\plus{2}&B&---&Y&Y&---&Y&---
\\

\midrule
\multicolumn{22}{c}{French Helicopters}\\
\midrule

Gazelle&
---&5&\plus{0}&\minus{1}&\arcminus{30}&4-8&
---&---&---&---&---&
Ballistic&\binarymultiply{A}{1}&\plus{0}&---&---&---&---&---&---&---
\\

\midrule
\multicolumn{22}{c}{Generic Helicopters\addedin{2A}{2A-helicopter-door-guns}{ and Door Guns}}\\
\midrule

Scout/Utility&
---&5&\plus{0}&\minus{2}&\arcminus{30}&4-8&
---&---&---&---&---&
---&\binarymultiply{A}{1}&\plus{2}&---&---&---&---&---&---&---
\\

Medium Lift&
---&7&\plus{0}&\plus{0}&\arcminus{60}&4-12&
---&---&---&---&---&
---&\binarymultiply{A}{1}&\plus{2}&---&---&---&---&---&---&---
\\

Heavy Lift&
---&8&\plus{1}&\plus{0}&\arcminus{30}&3-12&
---&---&---&---&---&
---&\binarymultiply{A}{1}&\plus{2}&---&---&---&---&---&---&---
\\

\tablerowaddedin{2A}{2A-helicopter-door-guns}{
Door LMG&
&&&&&&
7.62 mm LMG&LSD, RSD, or RD&10&1/1/1&\phantom{**}1/1**\phantom{}
}
\tablerowaddedin{2A}{2A-helicopter-door-guns}{
Door HMG&
&&&&&&
.50 cal HMG&LSD, RSD, or RD&10&1/1/1&\phantom{**}2/1**\phantom{}
}
\tablerowaddedin{2A}{2A-helicopter-door-guns}{
Door Minigun&
&&&&&&
7.62 mm Minigun&LSD, RSD, or RD&8&1/1/1&\phantom{**}2/2**\phantom{}
}

\bottomrule
\end{tabularx}
\begin{tablenote}{\linewidth}
\begin{itemize}

    \itemdeletedin{2A}{2A-helicopter-door-guns}{Some helicopters may have side-door gunners with LMGs. These can fire into the \arc{60}/\arc{90}/\arc{120}/\arc{150} arcs. They have 10 ammunition, hit rolls of 2/--/--, and damage ratings of 1/1**.}

    \itemdeletedin{2A}{2A-helicopter-door-guns}{The arc for a side-door gun (LSD/RSD) is \arc{60}/\arc{90}/\arc{120}/\arc{150}. The arc for a rear-door gun (RD) is \arcminus{60}.}

    \item Ballistic bomb systems have a \minus{1} modifier. Computed bomb systems have a \minus{2} modifier.

    \item The IRM suppression modifier is applied to IRM attack rolls.
    
    \item For attack helicopters, typical weapon loads are:
    \begin{itemize}
        \item AH-1J: Four RPs or two RPs and four TOW RGs. Additionally, may carry two AIM-9 IRMs or four Stinger IRM.
        \item AH-64: Four stations. Two may each carry four Helfire RGs or one RP. The other two may each carry one RP, one FT, or two Stinger IRMs.
        \item 500MD: Two stations. Each may carry one small RP, two TOW RGs, one GP, or two Stinger IRMs.
        \item OH-58D: Two stations. Each may carry one small RP, one GP with a caliber of 12.7 mm or less, or two Stinger IRMs.
        \item Mi-24: Four AT-2 Swatter RGs and four UB-16/UB-32 RPs
        \item Mi-24D: Four AT-2C Swatter RGs and four UB-16/UB-32 RPs
        \item Mi-24P: Four AT-6 Spiral RGs and four UB-16/UB-32 RPs. The AT-6 is air-to-air capable.
        \item Mi-28: Four stations. Each may carry four AT-6 Spiral RGs, one UB-16/UB-32 RP, one FT, or two SA-14 IRMs. At most four SA-14s may be carried.
        \item Gazelle: Four AS.11, HOT, or AT-3 RGs, or two RPs, or two GPs.
    \end{itemize}
\end{itemize}
\end{tablenote}

\end{fullwidthtable}

\end{landscape}
