%!TEX root = ../rules-working.tex
%LTeX: enabled=false


\begin{twocolumntablefloat}
\begin{twocolumntable}
\tablecaption{table:sighting-range-modifiers}{Relative Range Effects.}
\begin{tabularx}{0.75\linewidth}{lCCCCCCCC}
\toprule
Visual Range in Hexes&0--3&4--6&7--9&10--12&13--15&16--20&21--30&\plusafter{31}\\
Die Roll Modifier&\minus{2}&\minus{1}&\plus{0}&\plus{1}&\plus{2}&\plus{3}&\plus{5}&\plus{8}\\
\midrule
\\
\midrule
V.A.S. Range in Hexes&0--20&21--30&31--40&41--50&51--60&61--70&71--80&\plusafter{81}\\
Die Roll Modifier&\plus{0}&\plus{1}&\plus{2}&\plus{3}&\plus{4}&\plus{5}&\plus{6}&\plus{8}\\
\bottomrule
\end{tabularx}
\begin{tablenote}{0.75\linewidth}
\addedin{1C}{1C-apj-23-errata}{When looking for higher targets in daytime, count each 4 levels difference of altitude as 1 hex of range.}
\end{tablenote}
\end{twocolumntable}
\end{twocolumntablefloat}

\begin{twocolumntablefloat}
\begin{twocolumntable}
\tablecaption{table:sighting-position-modifiers}{Paint Scheme / Position / Weather Effects.}
\begin{tabularx}{0.75\linewidth}{lCCCCCC}
\toprule
Target Aircraft Position&Lower&Level&Higher&In Haze&In Stratus&Silh.\ by Cloud\\
\midrule
Silver          &\minus{2}&\minus{1}&\minus{1}&\plus{1}&\plus{3}&\minus{1}\\
Uncamouflaged   &\minus{1}&\plus{0}&\plus{0}&\plus{1}&\plus{3}&\minus{1}\\
Camouflaged     &\plus{1}&\plus{0}&\minus{1}&\plus{0}&\plus{3}&\minus{2}\\
Low Vis.\ Gray  &\plus{0}&\plus{1}&\plus{1}&\plus{2}&\plus{3}&\minus{1}\\
Aircraft Smoking&\minus{1}&\minus{2}&\minus{2}&NA&NA&NA\\
\bottomrule
\end{tabularx}
\end{twocolumntable}
\end{twocolumntablefloat}

\begin{onecolumntablefloat}
\begin{onecolumntable}
\tablecaption{table:sighting-additional-modifiers}{Additional Modifiers.}
\begin{tabularx}{\linewidth}{Xl}
\toprule
Number of aircraft searching:\\
\quad 1 or 2        &\plus{0}\\
\quad 3 or 4        &\minus{1}\\
\quad 5 to 8        &\minus{2}\\
\quad \plusafter{9} &\minus{3}\\
\textchangedin{2A}{2A-missile-sighting}{Tgt.\ is just launched Missile or SAM\asteriskmark}{Target is a missile with a normal boost or sustainer motor burning}&\minus{3}\\
\tablerowaddedin{2A}{2A-missile-sighting}{\textaddedin{2A}{2A-missile-sighting}{Target is a missile with a smokeless boost or sustainer motor burning}&\textaddedin{2A}{2A-missile-sighting}{\minus{1}}}
\tablerowaddedin{2A}{2A-missile-sighting}{\textaddedin{2A}{2A-missile-sighting}{Target is a missile just launched from a spotted aircraft}&\textaddedin{2A}{2A-missile-sighting}{\minus{2}}}
\textchangedin{2A}{2A-missile-sighting}{Tgt.\ is aircraft which just launched Missile\asteriskmark}{Target just launched a missile that was spotted}&\changedin{2A}{2A-missile-sighting}{\minus{3}}{\minus{2}}\\
\tablerowdeletedin{3A}{3A-searcher}{Tgt.\ is in all searcher's Restricted Arcs&\plus{2}}
\tablerowaddedin{3A}{3A-searcher}{Target in searcher's restricted arcs&\plus{2}}
Searcher is Looking out of Stratus&\plus{2}\\
Searcher is Veteran and Tactics Master&\minus{2}\\
Searcher is Veteran or Tactics Master&\minus{1}\\
Multi-crew aircraft Searching&\minus{1}\\
\changedin{2B}{2B-sighting-and-hud}{\changedin{1C}{1C-apj-25-qa/1C-apj-34-qa}{Hud Interface Technology used}{Searcher has HUD Interface Technology}}{Searcher has radar or IRSTS lock-on and HUD interface technology}&\minus{1}\\
Searcher has \changedin{2B}{2B-sighting-and-other-detections}{RWR indications}{detection or lock-on from radar, IRSTS, or RWR}&\minus{1}\\
Has Poor Eyes&\plus{1}\\
Has Good Eyes&\minus{1}\\
is Novice&\plus{1}\\
is Green&\plus{2}\\
Target using DDS Flare program&\changedin{1C}{1C-apj-36-errata}{\plus{PPL No.}}{\minus{PPL No.}}\\
\bottomrule
\end{tabularx}
\begin{tablenote}{\linewidth}
\deletedin{2A}{2A-missile-sighting}{\asteriskmark~does not apply to smokeless missiles.}

NOTE: All modifiers are cumulative. Disoriented or GLOC aircrew may not search.
\end{tablenote}
\end{onecolumntable}
\end{onecolumntablefloat}

