\begin{twocolumntablefloat}
\begin{twocolumntable}
\tablecaption{table:naval-critical-damage}{Naval Critical Damage.}
\begin{tabularx}{0.7\linewidth}{lP}
\toprule
First Die Roll&Damage Effects\\
\midrule
1&No effect\\
2&EWR disabled\\
3&Add-on FCR disabled\\
4&SAM launcher disabled\\
5&AAA site disabled\\
6&Fire starts. Roll a die at the end of each
turn. On a 1 the fire is put out. On a 10
it goes out of control and the ship is
considered to be crippled, unless
aluminum superstructure in which
case the fire is out of control on a
roll of 10, 9 or 8. (Fires aboard
Battleships and Carriers are not out
of control until a third 10 is rolled).

Regardless of the number of hits it
has taken, a vessel with an out of
control fire must roll for sinking at
the end of the game with a +5 die
modifier (cumulative with other
modifiers). 

Second or subsequent
Fire Starts results are treated
separately. However, once one
fire goes out of control, all other
fires are ignored.\\
7&Holed below the waterline: Regardless
of the number of hits it has taken, the
vessel must roll for sinking at the end of
the game. Each additional holed result
adds +1 to the sinking die roll.\\
8&No effect.\\
9&+1 hit (DO NOT roll for an additional
critical)\\
\plusafter{10}&Roll again on the table below.\\
\midrule
Second Die Roll&Damage Effects\\
\midrule
1&Explosive fire: An out of control fire starts (see above).\\
2&Fire starts in a magazine: A fire is started (see above) and a SAM or AAA is disabled (determine randomly).\\
3-4&+1 hit (DO NOT roll for an additional critical)\\
5&Jammer disabled (decrease jamming to 0).\\
6&Decoy launcher disabled (decrease decoy rating by one).\\
7-8&Fire control damaged: All SAMs are disabled; Jam and Decoy capability disabled.\\
9&Roll twice again on the Critical Damage Table.\\
\plusafter{10}& No effect.\\
\bottomrule
\end{tabularx}
\end{twocolumntable}

\medskip

\begin{tablenote}{0.7\linewidth}
Modify the first roll by +1 if the vessel has been damaged in a previous attack.

Modify the second roll by +1 for battleships and carriers.
\end{tablenote}
\end{twocolumntablefloat}
