\x{
\rulechapter{Positioning Game Counters}
}{
\rulechapter{Positions}
}

\x{
This chapter describes how the map hexes are utilized to position counters during play. The game map is used to show exactly where aircraft, missiles, and other units, are located in relationship to each other. Every game counter must be placed on the maps so that its position is clear to all players.
}{
This rule describes how the position of the game counters is defined with the help of the game map. The game map serves to show the exact location of aircraft, missiles, and other units. Every game counter must be placed on the map to make its location clear to all players. Aircraft and missiles in flight additionally have their position defined by their facing and altitude.
}

\x{
\section{Basic Elements of Position}
}{
\section{Elements of Position}
}
\label{rule:position}

\x{
\paragraph{Aircraft and Missile Counters.} 
Each aircraft and missile counter has a position which is defined by its map location, facing, and altitude.
}{
\paragraph{Aircraft and Missiles.} 
All aircraft and missiles have their position defined by their map location, facing, and altitude.
}

\begin{itemize}
    \item \itemparagraph{Map Location.} \x{
    With respect to the hexgrid, aircraft and missiles may be located wholly within a hex or on one of the lines between two hexes (commonly called a hexside). When located on a hexside, the game counter must face parallel to that hexside as illustrated \changedin{1C}{1C-figures}{below}{in Figure~\ref{figure:map-location}}. \deletedin{2B}{2B-range}{\addedin{1B}{1B-apj-23-errata}{When determining range when one or both aircraft are on hexsides, count only the full hexes between them. Take the shortest number of hexes.}}
    }{
    Aircraft and missile counters must be located wholly within a hex or on one of the lines between two hexes, commonly called a hex side. When located on a hex side, the game counter must face parallel to that hex side as illustrated in Figure~\ref{figure:map-location}.
    }

    \item \itemparagraph{Facing.} \x{
    This is the horizontal direction in which an aircraft or missile is flying. Use the silhouette on the counter to show facing by pointing the nose of the aircraft or missile in one of the twelve possible directions. Each direction differs by 30 degrees and has a corresponding compass heading associated with it (i.e., N = North, SSE = South South-East, etc.). In the scenario booklets, next to the map layout diagrams, there will be a compass arrow indicating which direction relative to the hexgrid that is North. \changedin{1C}{1C-figures}{See diagram below.}{See Figure~\ref{figure:facing}.}
    }{
    This is the horizontal direction in which an aircraft or missile is flying. Use the silhouette on the counter to show facing by pointing the nose of the aircraft or missile in one of the twelve possible directions shown in Figure~\ref{figure:facing}. Each direction differs by 30 degrees and has a corresponding compass heading associated with it (i.e., N is north, SSE is south-south-east, etc.). In the scenario booklets, next to the map layout diagrams, there will be a compass arrow indicating the direction of north relative to the hex grid.
    }

    \item \itemparagraph{Altitude.} \x{
    An aircraft or missile's altitude is kept track of on the aircraft's log sheet in terms of numbered levels. Each altitude level represents 1000 feet of height thus the number of a level corresponds to the altitude in thousands of feet (i.e., level 24 = 24,000 feet). Altitude levels are further grouped into named bands as described later in the rules. Aircraft and missile performance may vary in each altitude band as shown on an aircraft's data card and in the missile flight tables.
    }{
    An aircraft or missile's altitude is described in terms of numbered levels. Each altitude level corresponds to 1000 feet. Thus, the number of a level corresponds to the altitude in thousands of feet (i.e., level 24 corresponds to 24,000 feet). Altitude levels are further grouped into named bands as described later in the rules. Aircraft and missile performance may vary between altitude bands, as shown on the aircraft data cards and according to the missile flight rules.
    }

\end{itemize}

\input{figures/figure-map-location}

\input{figures/figure-facing}

\x{
\paragraph{Ground and Naval Unit Counters.}
Ground combat units and naval units have their positions defined simply by map location. They are always placed wholly within hexes and never on hexsides. Ground units never consider facing\addedin{1B}{1B-apj-23-errata}{ (except heavy AAA units under advanced rule 24.5)}, but large naval units must be faced in specific directions as for aircraft and missiles.
}{
\paragraph{Ground and Naval Units.}
Most ground and naval units have their position defined only by their map location, although some also use facing.
\begin{itemize}
    \item\itemparagraph{Map Location.} Ground and naval unit counters are always placed wholly within hexes and never on hex sides.
    
    \item\itemparagraph{Facing.} Heavy AAA units (see rule~\ref{rule:heavy-aaa-unit-facing}) and large naval units must be faced like aircraft and missiles. Other ground and naval units do not consider facing.

    \item\itemparagraph{Altitude.} Ground units and naval units on rivers or lakes are always at the altitude of terrain (see rule~\ref{rule:ground-terrain}). Naval units on seas or oceans are always at altitude level 0.
\end{itemize}
}

\Ax{
\section{Altitude Structure} 
Altitude levels are grouped into altitude bands, each several levels thick. Aircraft and missile performance can vary between the altitude bands, as the air density and temperature typically decrease with altitude. The altitude structure is shown in  Table~\ref{table:altitude-bands}.

\input {tables/table-altitude-bands}
}

\addedin{2B}{2B-range}{
\section{Range}
\label{rule:range}

Many rules depend on the range between two counters. Other than in the exceptional cases noted subsequently, the range in hexes between two counters is determined as using the following procedure.

\paragraph{Range Procedure.}
The range has two components, the horizontal and vertical ranges. They are determined and combined as follows:

\begin{itemize}
\item
The horizontal range is the length of the shortest route between the counters, measured in hexes. Moving between a hex side and one of the two adjacent hexes or four adjacent hex sides counts as one-half of a hex. After determining the number of hexes, including half hexes, round down.
\item
The vertical range is the difference in altitude levels divided by two and rounded down.
\item
The range is the sum of the horizontal and the vertical ranges.
\end{itemize}

\paragraph{Range Exceptions.}
The normal range procedure is not used in the following cases:
\begin{itemize}

\item
The range for air-to-air gun attacks is determined according to rule~\ref{rule:air-to-air-gunnery}.

\item
When attempting to search for a higher aircraft in daylight, the vertical range is the difference in altitude levels divided by \emph{four} and rounded down (see rule~\ref{rule:sighting-aircraft-and-missiles}).

\item
When considering positions of advantage, the horizontal range and vertical altitude difference are considered separately (see rule~\ref{rule:positions-of-advantage}).

\end{itemize}

}

\section{Stacking}

\x{
It is allowed, within the limits given below, for more than one counter to be in the same position on the map at the same time. When counters end up on top of other counters at the end of a turn, they are considered stacked. Stacking restrictions apply only al the end of a turn after all moves are completed.
}{
Counters are considered to be stacked when they occupy the same map location at the end of a game turn. Stacking is permitted within the limits given below. These limits only apply at the end of a game turn.
}

\x{

\paragraph{In The Air.}
Aircraft and missiles may freely fly through hexes or hexsides containing other game counters. They may freely stack on top of ground and naval units. They may freely stack with other aircraft and missiles that are not at the same altitude level. However, no more than two friendly aircraft may safely end up stacked together at the same altitude level (unless in Close Formation). Aircraft from opposing sides may not stack together at the same altitude safely. If these last two situations occur, you must check for collisions.

Exception: Up to four aircraft may be stacked together and may safely fly together at the same altitude while in a close formation. See advanced Rule 5.6.

\paragraph{On The Ground.}
No more than four ground unit counters may ever stack in the same ground terrain hex. No more than four small naval unit counters may ever stack in the same water, coastal, or river terrain hex. No more than two large naval unit counters may ever stack in the same water hex. When taxiing on the ground, up to four aircraft may move together in a stack.

}{

\paragraph{Aircraft.} Aircraft may stack with other aircraft, missiles, ground units, and naval units without limit. However, if aircraft are stacked at the same altitude level, there may be a risk of collision according to rule~\ref{rule:aircraft-collisions}.

\paragraph{Missiles.} Missiles may stack with other missiles, aircraft, ground units, and naval units without limit.

\paragraph{Ground Units.} No more than four ground unit counters may be stacked in the same ground terrain hex. No more than four aircraft may move together in a stack when taxiing on the ground.

\paragraph{Naval Units.} No more than four small naval unit counters may be stacked in the same water, coastal, or river terrain hex. No more than two large naval unit counters may be stacked in the same water hex.

}

\section{Aircraft Collisions}
\label{rule:aircraft-collisions}

\x{
Collisions are possible during the Flight Phase whenever an aircraft executes a head-on gun attack at range zero\addedin{1B}{1B-apj-36-errata}{ and the aircraft are at the same altitude}. Collisions are also possible at the end of \changedin{1B}{1B-apj-36-errata}{a game-turn}{the flight phase} in the following situations:
}{
\paragraph{Potential Collisions.}
Collisions are possible in the following situations:
}

\begin{itemize}
    \itemaddedin{2Z}{2Z-editing}{If an aircraft carries out a head-on gun attack at range 0 (see rule~\ref{rule:air-to-air-gunnery}) on a target at the same altitude.}
    \item \x{
    If an aircraft is stacked at the same altitude level with any enemy aircraft, and/or
    }{ 
    If an aircraft is stacked at the same altitude as an enemy aircraft after all aircraft have moved. However, aircraft in a tailing chain (see rule~\ref{rule:tailing-enemy-aircraft}) will not collide with each other.
    }
    \item \x{
    If an aircraft is stacked \addedin{2B}{2B-stacking}{at the same altitude }with two or more friendly aircraft and not in Close Formation.
    }{
    If more than two friendly aircraft are stacked at the same altitude after all aircraft have moved and they are not in a close formation (see rule~\ref{rule:close-formations}).
    }
    \itemaddedin{2B}{2B-stacking}{If more than four friendly aircraft are stacked at the same altitude after all aircraft have moved, even if they are in a close formation.}
\end{itemize}

\x{
\paragraph{Collision Resolution.} 
\addedin{2B}{2B-collisions}{Potential collisions resulting from head-on gun attacks are resolved immediately after the attack. Other potential collisions are resolved at the end of the flight phase.} For each potential collision, the player who last moved into the position must roll the die. On a roll of 1, his aircraft collides with one of the others. Determine which aircraft it collides with randomly. Both aircraft immediately roll on the 10 column of the Damage table to determine their damage.

\paragraph{Collision Exceptions:} There are \changedin{1B}{1B-apj-36-errata}{two}{three} exceptions to the Potential Collision rule.

\begin{itemize}
    \item An aircraft tailing another will not collide with that aircraft.
    \item An aircraft which is flying in close formation will not collide with other friendly aircraft in that formation.
    \itemaddedin{1B}{1B-apj-36-errata}{An aircraft that has already rolled for a collision for a head-on gun attack does not roll again for a collision \addedin{2B}{2B-collisions}{at the same position }again at the end of the flight phase.}
\end{itemize}

\notein{1B}{AWF: There is a clarification of the collision exceptions in the APJ 23 errata, but it seems to restate the original exceptions. I have not included it.}

}{
\paragraph{Collision Resolution Procedure.} 
If an aircraft carries out a head-on gun attack at range 0 and the same altitude, it immediately checks for a collision. The other cases are resolved at the end of the flight phase. Of the aircraft that could collide and which have not already checked for a collision at the same position (because of a head-on gun attack), the one that moved last checks for a collision. To check for a collision, roll the die. On a roll of 1, the aircraft collides with one of the others. Determine the other aircraft randomly. Both aircraft suffer damage from a weapon with an attack rating of 10 (see rule~\ref{rule:aircraft-damage-resolution}).
}

