%!TEX root = ./rules-working.tex
%LTeX: enabled=false

\addedin{3A}{3A-updating-adcs}{
\appendixchapter{Updating ADCs}
\label{appendix:updating-adcs}

ADCs for many aircraft have been published in {\AirSup}, {\AirStr}, {\TSOH}, and {\APJ}. Still more have been created by players interested in less common aircraft. 

Many of the values in these ADCs remains relevant under the current rules, but some need updating. This appendix describes how to identify and update earlier ADCs.


\section{Identifying ADC Versions}

First-edition ADCs can be identified in one of two ways:
\begin{itemize} 
\item They have the idle power labeled “Idle FP” or similar and the speedbrake row labeled “Spbr.\ FP” or similar. 
\item They have “v1” in the upper right, next to the text “Air Power”.
\end{itemize}

Second-edition ADCs can be identified in one of two ways:
\begin{itemize} 
\item They have the idle power labeled “Idle DP” or similar and the speedbrake row labeled “Spbr.\ DP” or similar. 
\item They have “v2” in the upper right, next to the text “Air Power”.
\end{itemize}

\addedin{3A}{3A-turn-drag}{
Third-edition ADCs can be identified in one way:
\begin{itemize} 
\item They have “v3” in the upper right, next to the text “Air Power”.
\end{itemize}
}

\section{Updating First-Edition ADCs}

To update first-edition ADCs for the second-edition rules:

\begin{itemize}

\itemaddedin{2A}{2A-idle}{Multiply the “Idle FP” values by two and interpret them as Idle DP values.}

For example, if the original idle value is 1.0 FP, the updated value is 2.0 DP.

\itemaddedin{2A}{2A-spbr}{Multiply the “Spbr.\ FP” values by two and interpret them as SPBR DP values.}

For example, if the original Spbr.\ value is 1.0 FP, the updated value is 2.0. DP

\itemaddedin{2B}{2B-roll-hfp-cost}{All displacement and lag rolls cost 1.0 HFP. Remove the roll HFP cost or replace it with 1.0.}

\end{itemize}

\silentlyaddedin{3A}{}{

Then apply the additional changes described below to update a second-edition ADC for the third-edition rules.

\section{Updating Second-Edition ADCs}

To update second-edition ADCs for the third-edition rules:

\begin{itemize}

    \itemaddedin{3A}{3A-turn-drag}{Take the original turn drag values, halve them, and add 0.5 for normal aircraft, 0.25 for LBR aircraft, and 0.75 for HBR aircraft. 

    For example, if the original turn drag is 2.0, the updated turn drag is 1.5 for normal aircraft, 1.25 for LBR aircraft, and 1.75 for HBR aircraft.}

\end{itemize}

}
}