\CX{

\begin{onecolumntablefloat}
\begin{onecolumntable}

\tablecaption{table:acceleration}{Accel/Decel Point Summary}

\begin{tabularx}{\linewidth}{C}
\toprule
Accel Point Summary\\
\midrule
\begin{itemize}
    \item Aircraft power = + Variable.
    \itemdeletedin{2A}{2A-steep-dives/2A-unloaded-dives}{Steep or Unloaded dive = +0.5 per level initially, then +1.0 per level.}
    \itemdeletedin{2A}{2A-steep-dives/2A-unloaded-dives}{Vert. dive = +1.0 per lvl.}
    \itemaddedin{2A}{2A-steep-dives/2A-unloaded-dives}{Steep, unloaded, and vertical  dive = 1.0 AP per level.}
\end{itemize}
\\
\midrule
Decel Point Summary\\
\midrule
\begin{itemize}
    \item Turning = Variable.
    \item Sust. climb = 0.5 per lvl.
    \item \changedin{2A}{2A-zoom-climbs}{Zooms = 1.0 per level initially, then 1.5 per lvl.}{Zoom climb = 1.0 DPs per level.}
    \item Vert. climb = \changedin{2A}{2A-vertical-climbs}{2.0}{1.5} per lvl.
    \item \changedin{2A}{2A-spbr}{Speed brake usage = 1.0 per 0.5 speed lost.}{Speedbrakes = up to the maximum value on the ADC.}
    \itemaddedin{2A}{2A-idle}{Idle power = the DP value on the ADC.}
    \item 1.0 if Idle or Normal Pwr.\ and above cruise speed\addedin{2A}{2A-cruise}{\ for the configuration}.
    \itemdeletedin{2A}{2A-sustained}{\changedin{1B}{1B-apj-23-errata}{Sustained turns and rolls 1.0 each, or 2.0 if HBR.}{Sustained turns 1.0 each or 2.0 if HBR.}}
    \itemaddedin{2A}{2A-sustained}{Sustained turns = 1.0 DP per 30 degree facing change in the second and subsequent facing changes (0.5 if LBR and 1.5 DP if HBR).}
    \itemaddedin{1B}{1B-apj-23-errata}{Sustained rolls 1.0 each}.
\end{itemize}
\\
\bottomrule
\end{tabularx}

\end{onecolumntable}
\end{onecolumntablefloat}

}{

\begin{twocolumntablefloat}[tp]
\begin{twocolumntable}

\tablecaption{table:acceleration}{AP and DP Summary}

\begin{tabularx}{0.9\linewidth}{lLl}
\toprule
Situation&Effect&Rule\\
\midrule
\addlinespace
Increase speed by 0.5&2.0 APs (1.5 if RA)&\mbox{\ref{rule:acceleration-and-deceleration-points}}\\
Increase speed by 0.5 at supersonic speeds&3.0 APs (2.0 if RA)&\mbox{\ref{rule:supersonic-speed-effects}\asteriskmark}\\
Decrease speed by 0.5&2.0 DPs&\mbox{\ref{rule:acceleration-and-deceleration-points}}\\
\addlinespace
\midrule
\addlinespace
Military or afterburner power&APs up to the maximum on the ADC as modified by damage, flame-outs, and altitude.&\mbox{\ref{rule:engine-thrust}}\\
VH altitude band (36--45)&Engine thrust is 2/3 normal (but not less than 0.5), unless the aircraft has HAE.&\mbox{\ref{rule:loss-of-thrust-with-altitude-and-speed}\asteriskmark}\\
EH or UH altitude bands (46+)&Engine thrust is 1/3 normal (but not less than 0.5), unless the aircraft has HAE.&\mbox{\ref{rule:loss-of-thrust-with-altitude-and-speed}\asteriskmark}\\
\addlinespace
Steep dives&1.0 AP for each level.&\mbox{\ref{rule:steep-dives}}\\
Unloaded dives&1.0 AP for each level.&\mbox{\ref{rule:unloaded-dives}}\\
Vertical dives&1.0 AP for each level.&\mbox{\ref{rule:vertical-dives}}\\
\addlinespace
\midrule
\addlinespace
Idle power&DPs according to the ADC.&\mbox{\ref{rule:engine-thrust}}\\
Idle power above the cruise speed&1.0 additional DP.&\mbox{\ref{rule:engine-thrust} \& \ref{rule:cruise-speeds}\asteriskmark}\\
Idle power at supersonic speeds&1.0 additional DP plus 2.0 additional DPs for each 0.5 of speed above high-transonic speed. (The 1.0 DP penalty for idle power above the cruise speed also applies.) Automatic flame-out.&\mbox{\ref{rule:supersonic-speed-effects}\asteriskmark \& \ref{rule:engine-flame-outs}\asteriskmark}\\
Normal power above the cruise speed&1.0 additional DP. Automatic flame-out if supersonic.&\mbox{\ref{rule:engine-thrust}\asteriskmark \& \ref{rule:engine-flame-outs}\asteriskmark}\\
Military power at supersonic speeds&1.5 DPs for each 0.5 of speed above high-transonic speed.&\mbox{\ref{rule:supersonic-speed-effects}\asteriskmark}\\
\addlinespace
Speedbrakes& DPs up to the maximum on the ADC.&\mbox{\ref{rule:speedbrakes}\asteriskmark}\\
Speedbrakes at supersonic speeds&Up to 1.0 additional DP.&\mbox{\ref{rule:supersonic-speed-effects}\asteriskmark}\\
\addlinespace
Transonic speeds or M1&DPs according to Table~\ref{table:transonic-and-supersonic-drag}.&\ref{rule:supersonic-speed-effects}\\
\addlinespace
Turning&DPs according to the ADC and the maximum turn rate.&\mbox{\ref{rule:turning}}\\
\RDY[3A-turn-drag]{Sustained turning&1.0 additional DP for each 30{\deg} of facing change after the first facing change (0.5 DPs if LBR and 1.5 DPs if HBR).&\mbox{\ref{rule:sustained-turning}\asteriskmark}}
Turning at supersonic speeds&\CY[3A-turn-drag]{1.0 additional DP for the game turn (0.0 DPs if GSSM and 2.0 DPs if PSSM)}{0.5 additional DPs for each turn declared (0.0 DPs if GSSM and 1.0 DPs if PSSM)}.&\mbox{\ref{rule:supersonic-speed-effects}\asteriskmark}\\
\addlinespace
Zoom climb&1.0 DPs for each level.&\mbox{\ref{rule:zoom-climbs}}\\
Sustained climb&0.5 DPs for each level up to the CC, then 1.0 DPs for each level.&\mbox{\ref{rule:sustained-climbs}}\\
Vertical climbs&2.0 DPs for each level.&\mbox{\ref{rule:vertical-climbs}}\\
\addlinespace
Rolling maneuvers&DPs for each roll according to the ADC.&\mbox{\ref{rule:rolling-maneuvers}}\\
\RDY[3A-rolls]{Sustained rolling&1.0 additional DP for each roll after the first.&\mbox{\ref{rule:sustained-rolling-maneuvers}}}
Rolling maneuvers at supersonic speeds&\CY[3A-rolls]{1.0 additional DP for the game turn (0.0 DPs if GSSM and 2.0 DPs if PSSM)}{1.0 additional DP per roll declared (0.0 DPs if GSSM and 1.0 DPs if PSSM)}&\mbox{\ref{rule:supersonic-speed-effects}\asteriskmark}\\
\addlinespace
VIFF maneuvers&DPs for each maneuver according to the maneuver.&\mbox{\ref{rule:viff-maneuvers}\asteriskmark}\\
\addlinespace
\bottomrule
\end{tabularx}
\begin{tablenote}{0.9\linewidth}
\asteriskmark~Advanced rule.
\end{tablenote}

\end{twocolumntable}
\end{twocolumntablefloat}

}
