\begin{onecolumntablefloat}[t]
\begin{onecolumntable}
\CX{
\tablecaption{table:system-loss}{Systems Loss.}
\begin{tabularx}{\linewidth}{cX}
\toprule
\multirow{2}{*}{\minitable{c}{Die\\Roll}}&
\multicolumn{1}{c}{Critical System Lost}\\
&\\
\midrule
1&Cockpit: Pilot Killed, remove aircraft from play.\\
2&Cockpit: Crewman killed. Lose multi-crew bonuses and lose radar and weapon technology. Bomb system = manual.\\
3&One engine permanently flamed out\\
4,5&Radar disabled. Lose all radar functions.\\
6,7&ECM disabled. Lose all ECM functions.\\
8&Weapons System disabled, aircraft may no longer attack. Jettison stores.\\
9&Internal guns and any gunpods disabled.\\
10&Technology disabled, lose all technology.\\
\bottomrule
\end{tabularx}
}{
\tablecaption{table:system-loss}{System Loss.}
\begin{tabularx}{\linewidth}{cX}
\toprule
\multirow{2}{*}{\minitable{c}{Die\\Roll}}&
\multicolumn{1}{c}{\multirow{2}{*}{Effects}}\\
&\\
\midrule
\addlinespace
1&Cockpit. The pilot is killed. Remove the aircraft from play.\\
\addlinespace
2&Cockpit. A non-pilot crewmember is killed. The aircraft lose multi-crew bonuses, loses radar and weapon technology (except boresight and auto-track modes), and is treated as having a manual bomb system.\\
\addlinespace
3&Engine. One engine is permanently flamed out\\
\addlinespace
4 or 5&Radar. The aircraft loses all radar functions.\\
\addlinespace
6 or 7&ECM. The aircraft loses all ECM functions.\\
\addlinespace
8&Weapons System. The aircraft may no longer attack and must jettison stores.\\
\addlinespace
9&Weapons System. The aircraft may no long use internal guns or gun pods.\\
\addlinespace
10&Technology. The aircraft loses all technology.\\
\addlinespace
\bottomrule
\end{tabularx}

}
\end{onecolumntable}
\end{onecolumntablefloat}