\begin{twocolumntablefloat}
\begin{twocolumntable}
\CX{

\tablecaption{table:stores-WR}{Aircraft Accessories}
\medskip
\begin{tabularx}{0.7\linewidth}{lcccL}
\toprule
\multicolumn{5}{c}{Weapon Racks “WR”}\\
\midrule
Type&Code&Weight&Load&Capacity\\
\midrule
Dual Rack&DR&100&1.0&2 weapons up to 1100 lbs.\ each.\\
Triple Rack&TR&100&1.0&3 Weapons up to 1100 lbs.\ each.\\
Multi.\ Rack&MR&200&2.0&4 to 6 weapons up to 800 lbs.\ each.\\
\midrule
\multicolumn{5}{c}{U.S. Special Racks}\\
\midrule
Missile-DR&MDR&100&1.0&Two Aim-9 Sidewinder missiles.\\
ARM-DR&ADR&200&2.0&Two AGM-45 Shrike ARMs.\\
\bottomrule
\end{tabularx}
\begin{tablenote}{0.7\linewidth}
\begin{enumerate}[nosep]
    \item All weapons on a rack must be identical.
    \item DRs and TRs may carry BB, BG, RP, RK, \& AGM-65 RG and RS weapons.
    \item MRs may carry any BB class weapons.
    \itemaddedin{1B}{1B-apj-23-errata}{F-8s may use MDRs to carry LAU-33 RPs.}
\end{enumerate}
\end{tablenote}

}{

\tablecaption{table:stores-WR}{Weapon Racks (WR)}
\medskip
\begin{tabularx}{0.7\linewidth}{llccL}
\toprule
\multicolumn{5}{c}{Weapon Racks}\\
\midrule
Type&Code&Weight&Load&Capacity\\
\midrule
Dual Rack&DR&100&1.0&2 weapons of up to 1100 lb each.\\
Triple Rack&TR&100&1.0&3 weapons of up to 1100 lb each.\\
Multiple Rack&MR&200&2.0&4 to 6 weapons of up to 800 lb each.\\
\midrule
\multicolumn{5}{c}{US Special Weapon Racks}\\
\midrule
Missile DR&MDR&100&1.0&Two AIM-9 Sidewinder missiles.\\
ARM DR&ADR&200&2.0&Two AGM-45 Shrike ARMs.\\
\bottomrule
\end{tabularx}
\begin{tablenote}{0.7\linewidth}
\begin{itemize}[nosep]
    \item All weapons on a rack must be identical.
    \item DRs and TRs may carry BB, BG, RP, RK, AGM-65 RG, and RS weapons.
    \item MRs may carry any BB class weapons.
    \itemaddedin{1B}{1B-apj-23-errata}{F-8s may use MDRs to carry LAU-33 RPs.}
\end{itemize}
\end{tablenote}

}
\end{twocolumntable}
\end{twocolumntablefloat}