\begin{onecolumntablefloat}[p]
\begin{onecolumntable}
\CX{
\tablecaption{table:supersonic-speed}{Transonic/Supersonic Speed}
\begin{tabularx}{1.0\linewidth}{Xccc}
\toprule
Alt. Band&Low Trans.&High Trans.&Mach One\\
\midrule
LO, ML&6.5&7.0&7.5\\
MH,HI&6.0&6.5&7.0\\
VH+&5.5&6.0&6.5\\
\bottomrule
\end{tabularx}
}{
\tablecaption{table:supersonic-speed}{Transonic and Supersonic Speeds}
\begin{tabularx}{1.0\linewidth}{XCCCC}
\toprule
Altitude Bands&Altitude Levels&Low-Transonic Speed&High-Transonic Speed&Mach One\\
\midrule
LO \& ML&\phantom{0}0--16&6.5&7.0&7.5\\
MH \& HI&17--35&6.0&6.5&7.0\\
VH+&36+&5.5&6.0&6.5\\
\bottomrule
\end{tabularx}
}
\end{onecolumntable}
\end{onecolumntablefloat}

\begin{onecolumntablefloat}[p]
\begin{onecolumntable}
\CX{
\tablecaption{table:transonic-and-supersonic-drag}{Transonic/Supersonic Drag Penalty}
\begin{tabularx}{1.0\linewidth}{Xccc}
\toprule
Aircraft Type&Low Trans.&High Trans.&Mach One\\
\midrule
LTD&0.0&0.5&1.0\\
NORMAL&0.5&1.0&1.5\\
HTD&1.0&1.5&2.0\\
\bottomrule
\end{tabularx}
}{
\tablecaption{table:transonic-and-supersonic-drag}{Transonic and Supersonic Drag}
\begin{tabularx}{1.0\linewidth}{LCCC}
\toprule
Aircraft Transonic Drag&Low-Transonic Speed&High-Transonic Speed&Mach One\\
\midrule
Low&0.0&0.5&1.0\\
Normal&0.5&1.0&1.5\\
High&1.0&1.5&2.0\\
\bottomrule
\end{tabularx}
}
\end{onecolumntable}
\end{onecolumntablefloat}

\begin{onecolumntablefloat}
\begin{onecolumntable}

\CX{
\tablecaption{table:supersonic-penalties}{Supersonic Penalties}
\begin{tabularx}{\linewidth}{X}
\toprule
\begin{itemize}[nosep]
    \item Add 1 prep to all maneuvers\deletedin{2A}{2A-snap}{ and snap turns}. Climb cap. = 2/3ds
    \item PSSM aircraft = $+2.0$ decel if any turns \changedin{1B}{1B-apj-23-errata}{or rolls}{and $+2.0$ decel if any rolls} done, and reduce maximum turn rate by one but not to less than HT.
    \item Normal aircraft = $+1.0$ decel if any turns \changedin{1B}{1B-apj-23-errata}{or rolls}{and $+1.0$ decel if any rolls} done.
    \item GSSM aircraft = No additional decel for turns or rolls.
    \item If in Mil.\ pwr., \changedin{1B}{1B-apj-23-errata}{$+1.0$}{$+1.5$} decel per 0.5 speed over High Transonic.
    \itemdeletedin{2A}{2A-idle/2A-supersonic-flame-out}{If in Normal pwer., $+2.0$ decel per 0.5 speed over High Transonic.}
    \itemdeletedin{2A}{2A-idle/2A-supersonic-flame-out}{If in Idle pwr., lose 0.5 more speed than listed on ADC.}
    \itemaddedin{2A}{2A-supersonic-flame-out}{If idle or military power selected, automatic flame-out.}
    \itemaddedin{2A}{2A-idle}{If all engines flamed-out, DPs for idle power from ADC, plus 1 DP for idle power at supersonic speed, plus 1 DP for idle power above cruise speed, plus 2 DP for each 0.5 of speed above high-transonic speed.}
    \item Takes 3.0 accel to gain 0.5 speed (2.0 if Rapid Accel aircraft).
\end{itemize}
\\
\bottomrule
\end{tabularx}
}{
\tablecaption{table:supersonic-penalties}{Supersonic Penalties}
\begin{tabularx}{\linewidth}{lX}
\toprule
\addlinespace
Idle power&All engines flame out.\\
\addlinespace
Normal power&All engines flame out.\\
\addlinespace
Turn rate&Reduced by one level for PSSM aircraft, but not to less than HT.\\
\addlinespace
Maneuvers&One additional preparatory HFP for all maneuvers. \\
\addlinespace
Climb capability&Reduced to {\twothirds}.\\
\addlinespace
\bottomrule
\end{tabularx}
}
\end{onecolumntable}
\end{onecolumntablefloat}
