\addedin{1B}{1B-apj-22-damage-tables}{

\begin{twocolumntable}
\x{

\tablecaption{table:optional-damage}{Optional Advanced Aircraft Damage}
\small
\begin{tabularx}{\linewidth}{lcX}
\toprule
\multirow{2}{*}{\minitable{c}{Die\\Roll}}&
\multirow{2}{*}{\minitable{c}{Hit\\Location}}&
Damage Effects to Aircraft and Notes\\
\\
\midrule
\multicolumn{3}{c}{Light Damage}\\
\midrule
&&$+1$ to all attacks, $+1$ to weapons launches.\\
1&Airframe&No supersonic speeds allowed.$^*$\\
2&Airframe&No BT or ET turn rate allowed.\\
3&Controls&Aircraft becomes Low roll rate and loses any high pitch rate capability.\\
4&Controls&Add one to prep-move requirements for all maneuvers.\\
5&Fuel&Lose 1 point of fuel per game turn, double bingo fuel amount. White Vapor emitted.\\
6&Avionics&Radar knocked out and lose all technology.\\
7&Avionics&Bombsight degraded to manual, gunsight = TT+1, HT+2, BT+3.\\
8&Weapons&Internal guns/gun pods disabled.\\
9&Engines&Afterburner disabled. No A/B power available.\\
10&Engines&Thrust loss; reduce Mil and A/B power numbers by 0.5. White Vapor emitted.\\
\midrule
\multicolumn{3}{c}{Heavy Damage}\\
\midrule
&&+2 to all attacks, +2 to weapon launches.\\
1,2&Airframe&No supersonic flight.$^*$ No BT or ET turns. Low roll rate. No high pitch rate. Add two to prep-move requirements for all maneuvers.\\
3&Controls&No rolling maneuvers allowed. Low roll rate. No high pitch rate. No ET turn rate.\\
4&Controls&Throttle jammed, no power changes. No HT, BT, or ET turn rate. No high pitch rate.\\
5&Fuel&Lose 2 points of fuel per turn. Bingo fuel tripled. Roll one die each turn after flight. On 1, a FUEL FIRE occurs (see description below). Aircraft becomes Smoker.\\
6,7&Avionics&Radar knocked out, all ECM knocked out, and lose all technology\\
8&Weapons&Guns disabled, no RHM/AHM/ARM/ASM missile or BS/RS weapon launches allowed.\\
9,10&Engines&Mil and A/B power halved. Roll for flame out whenever power setting changed. On 1, a Flame-Out occurs in one good engine. Aircraft becomes a smoker.\\
\midrule
\multicolumn{3}{c}{Critical Damage}\\
\midrule
&&$+3$ to all attacks, $+3$ to weapon launches. Roll once on the H table and once below.\\
1&Cockpit&Roll die again: 1, 2 = pilot killed (aircraft lost). 3, 4 = Crewman Killed (lose multi-crew functions). +5 = crew okay. Regardless, treat aircraft as having both an Airframe and Avionics hit as described immediately below.\\
2,3&Airframe&
No supersonic flight.$^*$ Low roll rate. No high pitch. No rolling maneuvers. EZ turns only. Aircraft must jettison external stores.\\
4,5,6&Avionics&Lose radar, all ECM, and all technology. No external pods work. Bombsight degrades to manual and gunsight = TT+1, HT+2, BT+3.\\
7&Fuel Fire**&Aircraft becomes Smoker. Lose 3 die rolls of fuel/turn. Jettison external stores. Roll die after each turn after flight. On 1 or 2 aircraft explodes, crew killed. On 9 or 10, fire goes out permanently but fuel leak still there.\\
8,9,10&Engines&A/B power lost. Mil power halved. Aircraft becomes Smoker. Roll for flame-outs twice per engine (1 or 2 = F.O. in this case). On subsequent turns, roll for flame-out whenever power seting is changed (1 = F.O.).\\
\bottomrule
\end{tabularx}
\begin{tablenote}{\linewidth}
\begin{itemize}
    \item The specifics of the damage rolled are not revealed to the enemy, but are recorded on paper. Visible signs of damage such as smoke or vapor, as indicated on the tables, must be told to the opponent.
    \item Jettison of Stores: Only the results calling for jettison of ordnance require an aircraft to do so. If this is the case, the aircraft must jettison enough stores to become CL configured. 
    \item If table indicates damage to a system or performance capability that the aircraft does not have, roll again. If the second roll also results in non-applicable damage, the hit is still recorded for the die roll penalties on combat and for possible progressive damage.
   \item[$^*$] If supersonic speeds exceeded in game turn, roll for progressive damage twice even if Damage Control was done.
   \item[$^{**}$] A Fuel Fire and risk of explosion can only be stopped by shutting down all engines. Declare this when doing damage control. Treat aircraft as flamed-out. When engines relit, A/B is disabled. Roll one die, on 1 to 4 fire resumes permanently; otherwise it stays out.
\end{itemize}
\end{tablenote}
\end{twocolumntable}

}{

\tablecaption{table:optional-damage}{Optional Aircraft Damage}
\small
\begin{tabularx}{\linewidth}{ccX}
\toprule
\multirow{2}{*}{\minitable{c}{Die\\Roll}}&
\multirow{2}{*}{\minitable{c}{Hit\\Location}}&
\multicolumn{1}{c}{\multirow{2}{*}{Effects}}
\\
\\
\midrule
\multicolumn{3}{c}{L or 2L Damage}\\
\midrule
&&$+1$ to all attacks and weapons launches.
\\
\cmidrule{3-3}
1&Airframe&
The aircraft must check twice for progressive damage at supersonic speeds.
\\
2&Airframe&
The aircraft may not use the BT or ET turn rates.
\\
3&Controls&
The aircraft becomes LRR and loses any HPR capability.
\\
4&Controls&
The aircraft requires one additional preparatory FP before all maneuvers.
\\
5&Fuel&
The aircraft loses 1 fuel point per game turn. 
Its bingo fuel level is doubled.
It emits a white vapor.
\\
6&Avionics&
The aircraft's radar and technology are disabled.
\\
7&Avionics&
The aircraft's bombsight is degraded to manual, and its gunsight to TT +1, HT +2, and BT +3.
\\
8&Weapons&
The aircraft can no longer use internal guns or gun pods.
\\
9&Engines&
The aircraft can no longer use afterburner power.
\\
10&Engines&
The aircraft's thrust at military and afterburner power is reduced by 0.5 AP.
It emits a white vapor.
\\
\midrule
\multicolumn{3}{c}{H Damage}\\
\midrule
&&+2 to all attacks and weapon launches.\\
\cmidrule{3-3}
1 or 2&Airframe&
The aircraft must check twice for progressive damage at supersonic speeds.
It may not use the BT or ET turn rates.
It becomes LRR and loses any HPR capability.
It requires two additional preparatory FP before all maneuvers.
\\
3&Controls&
The aircraft may not perform rolling maneuvers. 
It becomes LRR and loses any HPR capability.
It may not use the ET turn rate.
\\
4&Controls&
The aircraft's throttle jams. 
It may not change its power setting.
It may not use the HT, BT, or ET turn rates.
It loses any HPR capability.
\\
5&Fuel&
The aircraft loses 2 fuel points per game turn. 
Its bingo fuel level is tripled. 
It becomes a smoker.
The aircraft must check each game turn after flight for a fuel fire. On a die roll of $1-$, a fire starts.
\\
6 or 7&Avionics&
The aircraft's radar, ECM, and technology are disabled.
\\
8&Weapons&
The aircraft can no longer use internal guns or gun pods or launch RHM/AHM/ARM/ASM missiles or BS/RS weapons.
\\
9 or 10&Engines&
The aircraft's thrust at military and afterburner power is reduced to one-half the normal values.
It must check for a flame-out each time it changes its power setting. On a $1-$, one good engine flames out.
The aircraft becomes a smoker.
\\
\midrule
\multicolumn{3}{c}{C Damage}\\
\midrule
&&Roll once on the H table and once on the C table.\\
&&$+3$ to all attacks and weapon launches.\\
\cmidrule{3-3}
1&Cockpit&
Roll the die again: 1 or 2 = pilot killed (equivalent to K). 3 or 4 = crewmember killed (lose multi-crew functions). $5+$ = crew unhurt. 
Treat the aircraft as having both the airframe and avionics results described immediately below.
\\
2 or 3&Airframe&
The aircraft must check twice for progressive damage at supersonic speeds.
It may only use the EZ turn rate.
It becomes LRR and loses any HPR capability.
It may not perform rolling maneuvers. 
It must jettison stores.
\\
4 to 6&Avionics&
The aircraft's radar, ECM, technology, and all external pods are disabled.
Its bombsight is degraded to manual, and its gunsight to TT +1, HT +2, and BT +3.
\\
7&Fuel&
The aircraft suffers a fuel fire.
It loses 3 fuel points per game turn. 
%Its bingo fuel level is quadrupled. 
It becomes a smoker.
It must jettison stores.
%Roll die after each turn after flight. On 1 or 2 aircraft explodes, crew killed. On 9 or 10, the fire goes out permanently, but the fuel leak is still there.
\\
8 to 10&Engines&
The aircraft loses afterburner power, and its thrust at military power is reduced to one-half the normal value.
It becomes a smoker.
It immediately checks twice per engine for a flame-out on a $2-$. On subsequent turns, if it changes its power setting, each engine flames out on a $1-$.
\\
\bottomrule
\end{tabularx}

}

\end{twocolumntable}

}
