\begin{onecolumntable}[p]
\x{
\tablecaption{table:supersonic-speed}{Transonic/Supersonic Speed}
\begin{tabularx}{1.0\linewidth}{Xccc}
\toprule
Alt. Band&Low Trans.&High Trans.&Mach One\\
\midrule
LO, ML&6.5&7.0&7.5\\
MH,HI&6.0&6.5&7.0\\
VH+&5.5&6.0&6.5\\
\bottomrule
\end{tabularx}
}{
\tablecaption{table:supersonic-speed}{Transonic and Supersonic Speeds}
\begin{tabularx}{1.0\linewidth}{XCCCC}
\toprule
Altitude Bands&Altitude Levels&Low-Transonic Speed&High-Transonic Speed&Mach One\\
\midrule
LO \& ML&0--16&6.5&7.0&7.5\\
MH \& HI&17--35&6.0&6.5&7.0\\
VH+&36+&5.5&6.0&6.5\\
\bottomrule
\end{tabularx}
}
\end{onecolumntable}

\begin{onecolumntable}[p]
\x{
\tablecaption{table:transonic-and-supersonic-drag}{Transonic/Supersonic Drag Penalty}
\begin{tabularx}{1.0\linewidth}{Xccc}
\toprule
Aircraft Type&Low Trans.&High Trans.&Mach One\\
\midrule
LTD&0.0&0.5&1.0\\
NORMAL&0.5&1.0&1.5\\
HTD&1.0&1.5&2.0\\
\bottomrule
\end{tabularx}
}{
\tablecaption{table:transonic-and-supersonic-drag}{Transonic and Supersonic Drag}
\begin{tabularx}{1.0\linewidth}{LCCC}
\toprule
Aircraft Transonic Drag&Low-Transonic Speed&High-Transonic Speed&Mach One\\
\midrule
Low&0.0&0.5&1.0\\
Normal&0.5&1.0&1.5\\
High&1.0&1.5&2.0\\
\bottomrule
\end{tabularx}
}
\end{onecolumntable}

\begin{onecolumntable}

\x{
\tablecaption{table:supersonic-penalties}{Supersonic Penalties}
\begin{tabularx}{\linewidth}{X}
\toprule
\begin{itemize}[nosep]
    \item Add 1 prep to all maneuvers\deletedin{2A}{2A-snap}{ and snap turns}. Climb cap. = 2/3ds
    \item PSSM aircraft = $+2.0$ decel if any turns \changedin{1B}{1B-apj-23-errata}{or rolls}{and $+2.0$ decel if any rolls} done, and reduce maximum turn rate by one but not to less than HT.
    \item Normal aircraft = $+1.0$ decel if any turns \changedin{1B}{1B-apj-23-errata}{or rolls}{and $+1.0$ decel if any rolls} done.
    \item GSSM aircraft = No additional decel for turns or rolls.
    \item If in Mil.\ pwr., \changedin{1B}{1B-apj-23-errata}{$+1.0$}{$+1.5$} decel per 0.5 speed over High Transonic.
    \itemdeletedin{2A}{2A-idle and 2A-supersonic-flamed-out}{If in Normal pwer., $+2.0$ decel per 0.5 speed over High Transonic.}
    \itemdeletedin{2A}{2A-idle and 2A-supersonic-flamed-out}{If in Idle pwr., lose 0.5 more speed than listed on ADC.}
    \itemaddedin{2A}{2A-supersonic-flamed-out}{If idle or military power selected, automatic flame-out.}
    \itemaddedin{2A}{2A-idle}{If all engines flamed-out, DPs for idle power from ADC, plus 1 DP for idle power at supersonic speed, plus 1 DP for idle power above cruise speed, plus 2 DP for each 0.5 of speed above high-transonic speed.}
    \item Takes 3.0 accel to gain 0.5 speed (2.0 if Rapid Accel aircraft).
\end{itemize}
\\
\bottomrule
\end{tabularx}
}{
\tablecaption{table:supersonic-penalties}{Supersonic Effects}
\begin{tabularx}{\linewidth}{X}
\toprule
\begin{itemize}
    
    \item Increasing speed by 0.5 requires 3.0 APs (2.0 APs if RA).

    \item If idle or normal power is used, the engines automatically flame out.
    \item If all engines are flamed out, the aircraft receives DPs for idle power from the ADC, plus 1.0 DPs for idle power at supersonic speed, plus 1.0 DPs for idle power above cruise speed, plus 2.0 DPs for each 0.5 of speed above high-transonic speed.
    \item If military power is used, the aircraft receives 1.5 DPs per 0.5 speed over high-transonic speed.

    \item PSSM aircraft have their maximum turn rate reduced by one level, but not to less than HT.
    \item If the aircraft uses turning flight, it receives 1.0 additional DPs (0.0 DPs if GSSM and 2.0 DPs if PSSM) per game turn
    \item The aircraft's climb capability is reduced to {\twothirds}.
    \item The aircraft requires one additional preparatory HFP for all maneuvers. 
    \item If the aircraft executes any rolling maneuvers, it receives 1.0 additional DP (0.0 DPs if GSSM and 2.0 DPs if PSSM) per game turn.

\end{itemize}
\\
\bottomrule
\end{tabularx}
}

\end{onecolumntable}
