\begin{onecolumntable}

\x{

\tablecaption{table:acceleration}{Accel/Decel Point Summary}

\begin{tabularx}{\linewidth}{C}
\toprule
Accel Point Summary\\
\midrule
\begin{itemize}
    \item Aircraft power = + Variable.
    \itemdeletedin{2A}{2A-steep-dives and 2A-unloaded-dives}{Steep or Unloaded dive = +0.5 per level initially, then +1.0 per level.}
    \itemdeletedin{2A}{2A-steep-dives and 2A-unloaded-dives}{Vert. dive = +1.0 per lvl.}
    \itemaddedin{2A}{2A-steep-dives and 2A-unloaded-dives}{Steep, unloaded, and vertical  dive = 1.0 AP per level.}
\end{itemize}
\\
\midrule
Decel Point Summary\\
\midrule
\begin{itemize}
    \item Turning = Variable.
    \item Sust. climb = 0.5 per lvl.
    \item \changedin{2A}{2A-zoom-climbs}{Zooms = 1.0 per level initially, then 1.5 per lvl.}{Zoom climb = 1.0 DPs per level.}
    \item Vert. climb = 2.0 per lvl.
    \item \changedin{2A}{2A-spbr}{Speed brake usage = 1.0 per 0.5 speed lost.}{Speedbrakes = up to the maximum value on the ADC.}
    \itemaddedin{2A}{2A-idle}{Idle power = the DP value on the ADC.}
    \item 1.0 if Idle or Normal Pwr.\ and above cruise speed\addedin{2A}{2A-cruise}{\ for the configuration}.
    \itemdeletedin{2A}{2A-sustained}{\changedin{1B}{1B-apj-23-errata}{Sustained turns and rolls 1.0 each, or 2.0 if HBR.}{Sustained turns 1.0 each or 2.0 if HBR.}}
    \itemaddedin{2A}{2A-sustained}{Sustained turns = 1.0 DP per 30 degree facing change in the second and subsequent facing changes (0.5 if LBR and 1.5 DP if HBR).}
    \itemaddedin{1B}{1B-apj-23-errata}{Sustained rolls 1.0 each}.
\end{itemize}
\\
\bottomrule
\end{tabularx}

}{
\tablecaption{table:acceleration}{AP/DP Summary}

\begin{tabularx}{\linewidth}{C}
\toprule
AP Summary\\
\midrule
\begin{itemize}
    \item Military or afterburner power = APs up to the limit on the ADC and modified by damage and altitude.
    \item Steep, unloaded, and vertical dive = 1.0 AP per level.
\end{itemize}
\\
\midrule
DP Summary\\
\midrule
\begin{itemize}
    \item Turning = DPs according to the ADC and the maximum turn rate.
    \item Sustained climb = 0.5 DPs per level up to the sustained climb limit.
    \item Zoom climb = 1.0 DP per level.
    \item Vertical climb = 2.0 per level.
    \item Speedbrakes = DPs up to the maximum value on the ADC.
    \item Idle power = DPs according to the ADC.
    \item Idle or normal power above the cruise speed for the configuration = 1.0 DP
    \item Sustained turns = 1.0 DP for each 30{\deg} facing change in the second and subsequent facing changes (0.5 DP if LBR and 1.5 DP if HBR).
    \item Sustained rolls = 1.0 DP each for the second and subsequent rolls.
\end{itemize}
\\
\bottomrule
\end{tabularx}

}
\end{onecolumntable}
