\begin{onecolumntable}
\tablecaption{table:crew-quality-generation}{Pilot/Crew Generation}
\small
\begin{tabularx}{\linewidth}{l*{5}{c}}
\toprule
&\multicolumn{5}{c}{National Training Standard}\\
\cmidrule(){2-6}
Quality&Excellent&Good&Average&Limited&Poor\\
\midrule
Veteran&1--3&1--2&1&1&NA\\
Regular&4--8&3--7&2--6&2--4&1--4\\
Novice&9--10&8--9&7--9&5--8&5--7\\
Green&NA&10&10&9--10&8--10\\
\bottomrule
\end{tabularx}
\begin{tablenote}{\linewidth}
\begin{itemize}[nosep]
    \item Roll one die per aircrew; reference training standard and roll to find crew quality at left. Example; die roll “6” under Good = Regular.
\end{itemize}
\end{tablenote}

\vspace{\floatsep}

\tablecaption{table:crew-attributes-generation}{Aircrew Attribute Determination}
\small
\begin{tabularx}{\linewidth}{lCCC}
\toprule
Attr. Level&Eyesight&Fitness&Confidence\\
\midrule
Excellent&1--2&1--3&1--2\\
Average&3--9&4--8&3--8\\
Poor&10&9--10&9--10\\
\bottomrule
\end{tabularx}
\begin{tablenote}{\linewidth}
\begin{itemize}[nosep]
    \item Roll once per attribute per aircrew; cross reference as above to find level of attribute (either excellent, average, or poor).
    \item Excellent Eyesight = $-1$ and Poor Eyesight = $+1$ for sighting die rolls.
    \item Excellent Fitness = $+1$ and Poor Fitness = $-1$ for GLOC and Post-Egress Fate rolls.
    \item Excellent Confidence = $+1$ and Poor Confidence = $-1$ to initiative, Departure and Post-Egress Fate die rolls.
\end{itemize}
\end{tablenote}

\vspace{\floatsep}

\tablecaption{table:crew-special-characteristics-generation}{Aircrew Special Characteristics Determination}
\small
\begin{tabularx}{\linewidth}{lCCC}
\toprule
\minitable{c}{Crew\\Quality}&\minitable{c}{Sierra\\Hotel}&\minitable{c}{Tactics\\Master}&\minitable{c}{Combat\\Hero}\\
\midrule
Veteran&1&1--3&1--2\\
Regular&1&1--2&1\\
Novice&1&1&NA\\
\bottomrule
\end{tabularx}
\begin{tablenote}{\linewidth}
\begin{itemize}[nosep]
    \item Roll once per characteristic per Veteran, Regular, and Novice aircrew. Result <= to above number gives them the characteristic.
    \item Tactics master acquisition modifiers = $-1$ if Training Standard = Excellent; $+1$ if Training Standard = Limited or Poor.
\end{itemize}
\end{tablenote}

\end{onecolumntable}