\begin{TABLE}

\TABLECAPTION{table:missile-flight-rules}{Missile Flight Rules}

\small
\begin{tabularx}{\linewidth}{X}
\toprule\\
\multicolumn{1}{c}{FP Costs}\\
\begin{enumerate}
    \item One FP to move forward one hex/hexside.
    \item One FP to climb 1 or 2 Altitude levels.
    \item Free lost of 1 level per hex entered.
    \item One FP to dive 2 or 3 Altitude levels.
    \item Once per turn may dive 1 level with 1 FP.
    \item One FP to\deletedin{2A}{2A-snap}{ Snap-turn or} Slide, 0 FP to vert.\ roll.
\end{enumerate}\\

\multicolumn{1}{c}{Manuever Limits}\\

\begin{enumerate}
    \itemdeletedin{2A}{2A-snap}{One Snap-turn allowed during entire flight.}
    \item Only Slide and Vertical roll maneuvers allowed.
    \item Normally, 1 Vert.\ roll allowed in entire flight except anytime target performs one, missiles may in next move.
    \itemdeletedin{2A}{2A-snap}{If Snap-turn first action other than forward flight after missile arms, no prep-move required.}
    \itemdeletedin{2A}{2A-snap}{If missile turns, or switches between climbs and dives before Snap-turning, normal prep-move must be met.}
\end{enumerate}\\

\multicolumn{1}{c}{Flight Restrictions}\\

\begin{enumerate}
    \item Missiles may climb and dive in same game turn, and some may do both in same proportional move.
    \item Vertical roll allowed when msl.\ expends 2 or more FPs while climbing or diving in same position.
    \item \changedin{1B}{1B-apj-23-errata and 1B-apj-24-play-aids}{If turn ability is not BT/2, ET/2, or ET/3 then missile is limited to switching between climbs and dives. Such missiles may do either in proportional move but not both. Before changing between the two, missile must spend 1 proportional move in level flt.}{If turn ability is not BT/2, ET/2, or ET/3 then missile is limited in switching between climbs and dives.  Such missiles must expend FPs equal to {\onethird} their speed (round up) before expending VFPs to change altitude in the opposite direction. If the turn ability is BT/2 or better, then only FPs equal to $1/10$ their speed (round up) need be spent in level flight before switch directions.}
    \item Missiles may never dive if already below their target.
    \item Missiles may only climb if already above their target if;
    \begin{enumerate}
        \item They are CG SAMs in boost phase.
        \item They are TVM SAMs of MCG missiles.
    \end{enumerate}
\end{enumerate}\\

\multicolumn{1}{c}{Missile Speed Changes}\\

\begin{enumerate}
    \item If missile gained altitude over turn, $-1$ to speed for each set of alt.\ levels climbed equal to half or less of missile's speed.
    \item If missile lost altitude over turn, $+1$ to speed for each set of alt.\ levels dived equal to half or more of missile's speed.
    \itemaddedin{1B}{1B-apj-23-errata and 1B-apj-24-play-aids}{If missile changed facing by turning, $-1$ to speed for each 30 degrees of facing change done.}
\end{enumerate}\\

\multicolumn{1}{c}{Missile Speed Determination}\\

\begin{enumerate}
    \item Air to air first turn = $(\textrm{Base} + \textrm{aircraft}) \times \textrm{Speed Att.\ Factor}$. \addedin{1B}{1B-apj-23-errata and 1B-apj-24-play-aids}{If the missile has a sustainer, use the factor for the altitude band two above the one in which the missile was launched.}
    \item SAM first turn = listed Base Speed.
    \item Subsequent turns = $(\textrm{Previous} \pm \textrm{changes})  \times \textrm{Speed Att.\ factor}$.
    \item \changedin{1B}{1B-apj-23-errata and 1B-apj-24-play-aids}{If sustainer motor in effect}{On subsequent turns, if the sustainer motor is in effect}, Speed Att.\ Factor = 1.0.
\end{enumerate}\\

\bottomrule
\end{tabularx}

\end{TABLE}
