\appendixchapter{Updating ADCs}

ADCs for many aircraft have been published in {\AirSup}, {\AirStr}, {\TSOH}, and {\APJ}. Still more have been created by players interested in less common aircraft. 

Many of the values in these ADCs remains relevant under the current rules, but some need updating. This appendix describes how to identify and update earlier ADCs.


\section{Identifying ADC Versions}

First-edition ADCs can be identified in one of two ways:
\begin{itemize} 
\item They have the idle power labeled “Idle FP” or similar and the speedbrake row labeled “Spbr.\ FP” or similar. 
\item They have “v1” in the upper right, next to the text “Air Power”.
\end{itemize}

Second-edition ADCs can be identified in one of two ways:
\begin{itemize} 
\item They have the idle power labeled “Idle DP” or similar and the speedbrake row labeled “Spbr.\ DP” or similar. 
\item They have “v2” in the upper right, next to the text “Air Power”.
\end{itemize}

Version 3 ADCs can be identified in one way:
\begin{itemize}
\item They have “v3” in the upper right, next to the text “Air Power”.
\end{itemize}

\section{Updating ADCs}

\subsection{Idle and Speedbrake DPs}

To update first-edition ADCs to the current rules:

\begin{itemize}

\item Multiply the “Idle FP” values by two and interpret them as Idle DP values.
\item Multiply the “Spbr.\ FP” values by two and interpret them as SPBR DP values.

\end{itemize}

For example, if the original idle FP values are 0.5, 0.5, and 1.0 (for CL, 1/2, and DT configurations), the updated values are 1.0, 1.0, and 2.0.

Second-edition and version 3 ADCs do not need updating in this way.

\AY{
\subsection{Turn Drag}

To update first-edition and second-edition ADCs to the current rules:

\begin{itemize}

\item Take the original turn drag values, halve them, and add 0.5 for normal aircraft, 0.25 for LBR aircraft, and 0.75 for HBR aircraft. 

\end{itemize}

For example, if the original turn drag is 2.0, the updated turn drag is 1.5 for normal aircraft, 1.25 for LBR aircraft, and 1.75 for HBR aircraft.

Version 3 ADCs do not need updating in this way.


}