\rulechapter{Aircraft Flight}

\input{tables/table-aircraft-flight-rules}

\x{
An aircraft is flown, in game turns, by moving it horizontally across the game maps and tracking its changes in speed and altitude on an Aircraft Log Sheet.\addedin{1C}{1C-tables}{\ The aircraft flight rules are summarized in Table~\ref{table:aircraft-flight-rules-summary} and detailed in the following chapters.} An aircraft is allowed to move and make changes in altitude by expending Flight Points (FPs). Generally, one flight point moves the aircraft one hex on the map, or up or down one or more levels of altitude. The speed of an aircraft determines how many flight points it has each turn.
}{
An aircraft is flown, in game turns, by moving it horizontally across the game maps and tracking its changes in speed and altitude on an aircraft log sheet. The aircraft flight rules are summarized in Table~\ref{table:aircraft-flight-rules-summary} and detailed in the following chapters. An aircraft can move and change altitude by expending Flight Points (FPs). Generally, one flight point moves the aircraft one hex on the game map, or up or down one or more altitude levels. The speed of an aircraft determines how many flight points it has each turn.
}

\x{
\section{The Aircraft Log}
}{
\section{Aircraft Log Sheets}
}
\label{rule:aircraft-log-sheets}

\x{
An aircraft log sheet is used to record the starting speed and altitude of an aircraft each turn. Different actions an aircraft may take will affect its speed and altitude from one turn to the next. The different lines on the log provide spaces to record these actions and to calculate resultant changes in speed and altitude. A pad of generic aircraft log sheets is provided with each game. One log should be used for each aircraft in play. A log sheet is divided into 15 columns, enough to record 15 turns of play (about the average length of a game). If you run short you may make copies of the sheets. Where applicable, the various rules sections which follow will provide additional information on using the log sheet.
}{
An aircraft log sheet is used to record an aircraft's starting speed and altitude each game turn. One log sheet should be used for each aircraft in play. An aircraft's actions in a game turn affect its speed and altitude from one turn to the next. The different lines on the log sheet provide spaces to record these actions and calculate resultant speed and altitude changes. A log sheet is divided into 15 columns and covers 15 game turns of play, about the average length of a game. Where applicable, the detailed rules that follow will provide additional information on using the log sheet.
}

\paragraph{\protect\x{Set Up Information}{Set-Up}.} \x{
When a scenario is set up, each aircraft must be given a start altitude level, and speed. These are noted in lines 1 and 2 of column one of the log sheets. Any scenarios provided in the game will establish these values in their set up instructions. If you create your own scenarios, you will have to decide on these values yourself. Once aircraft and any other required counters are placed on the maps, and the start speeds and altitudes are noted, play may commence.
}{
Each aircraft must be given a start altitude level, and speed at the start of a scenario. These are noted in the first column of the log sheets. The scenarios provided in the game will establish these values in their set-up instructions. If you create your own scenarios, you will have to decide on these yourself. Once aircraft and any other required counters have been placed on the maps and the start speeds and altitudes have been noted, play may commence.
}

\paragraph{Using The Log Sheet.} 
\x{
As an aircraft is maneuvered, it might change altitude and/or accumulate accel and decel points. The accumulation of accel and/or decel Points may cause changes in the aircraft's speed so spaces are provided on the log to record these points. Spaces are also provided to calculate changes in an aircraft's speed and/or altitude. The final altitude and speed at the end of one game turn becomes the new start speed and altitude for the next game turn.
}{
As an aircraft maneuvers, it can change altitude (see rule~\ref{rule:changing-aircraft-altitude}) It can also accumulate acceleration and deceleration points (APs and DPs), and these may cause changes in the aircraft's speed (see rule~\ref{rule:changing-aircraft-speed}). Spaces are provided in the log to record APs and DPs and to calculate changes in an aircraft's speed and altitude. The final altitude and speed at the end of one game turn becomes the new start speed and altitude for the next.
}

\section{Flight Points}
\label{rule:flight-points}

\x{
The speed of an aircraft is always expressed in terms of Flight Points (FPs), each of which equates to 100 mph of speed. An aircraft with a speed of 5.5 would be traveling at 550 mph. Flight Points are expended to move an aircraft across the game map and/or to change altitude.  An aircraft must expend all whole FPs available to it each game turn. Any unused 0.5 FP remaining can be ignored. An FP may be either a horizontal FP or a vertical FP depending on how it is used.
}{
The speed of an aircraft is expressed in terms of flight points (FPs), each of which equates to 100 mph of speed. An aircraft with a speed of 5.5 would be traveling at 550 mph. FPs are expended to move an aircraft across the game map and to change altitude. An aircraft must expend all whole FPs available to it each game turn. Any remaining unused half FP can be ignored. An FP may be either a horizontal FP (HFP) or a vertical FP (VFP), depending on how it is used. The mixture of HFPs and VFPs depends on the aircraft's flight type.
}

\paragraph{Horizontal Flight Points.} 
\x{
FPs expended to move horizontally across the map are called Horizontal Flight Points (HFPs). One HFP is spent to move an aircraft forward one hex or hexside. It may only fly onto a hexside if it faces parallel to that hexside (see 3.1).
}{
An aircraft expends horizontal FPs (HFPs) to move horizontally across the map. Each HFP moves the aircraft forward one hex or hex side. An aircraft may only fly onto a hex side if it faces parallel to that hex side (see rule~\ref{rule:position}).
}

\paragraph{Vertical Flight Points.} 
\x{
FPs expended to gain or lose altitude levels are called Vertical Flight Points (VFPs). The amount of altitude levels gained or lost with the expenditure of each VFP varies with the exact type of climb or dive in use.  The amount of FPs that may be VFPs in a game turn also depends on the exact type of climbing or diving flight chosen.

Note: To clarify; both HFPs and VFPs will be available to be expended within a single game turn only when an aircraft elects to climb or dive during its move.
}{
An aircraft expends vertical FPs (VFPs) to gain or lose altitude levels. The amount of altitude levels gained or lost with the expenditure of each VFP varies with the exact type of climb or dive in use. The number of FPs that may be VFPs in a game turn also depends on the type of climbing or diving flight chosen. VFPs will only be available to an aircraft if it elects climbing or diving flight during its move.
}

\addedin{1C}{1C-tables}{
    \input{tables/table-fractions}
}

\x{
\addedin{1B}{1B-apj-36-errata}{When an expenditure of $2/3$ of an aircraft's speed (or FPs) is required, use the $2/3$ entry from \changedin{1C}{1C-tables}{the chart}{Table~\ref{table:fractions}}, not $2 \times 1/3$. Drop all fractions read off the chart.}
}{
\paragraph{Fractions of Speed or FPs.} Some rules require the expenditure of FPs equal to {\onethird} or {\twothirds} of an aircraft's speed or total FPs. Use Table~\ref{table:fractions} to determine these fractions and round as directed by the rule. Where the rules call for {\twothirds}, use the value in the table, not twice {\onethird}.
}

\x{
\addedin{1B}{1B-apj-36-errata}{\paragraph{Speed Requirements Split Between Game Turns.} Some game activities, like ground-attack aiming, are measured in fractions of an aircraft's speed. If such a period extends between two game-turns, the highest speed in either game-turn is used to compute the completion requirements in the second turn.}
}{
\paragraph{Speed Requirements Split Between Game Turns.} Some game activities, like ground-attack aiming, are measured in fractions of an aircraft's speed. If such a period extends between two game turns, the highest speed in either game turn is used to compute the requirements in the second game turn.
}

\section{Types of Flight}

\x{
At the start of an aircraft's move, it must commit itself to one of the following three general types of flight: Level, Climbing or Diving. Write the appropriate code in the flight type line of the aircraft log. This represents the aircraft committing its nose to remain level, to move up, or to move down for the game turn. The flight type may not be changed until the next turn. Only one type of flight may be performed each turn and some restrictions may apply when switching from one type to another between turns.
}{
At the start of an aircraft's move, it must select one of the following three general types of flight: level, climbing, or diving. This represents the aircraft committing its nose to remain level, to move up, or to move down for the game turn. Write the appropriate code in the flight type line of the aircraft log. Once selected, aircraft may not change flight type until the next game turn. Aircraft may perform only one type of flight each game turn, and restrictions may apply when they switch to another type in the next game turn.
}

\paragraph{Level Flight.} 
\x{
In level flight, all FPs must be expended as HFPs (for example, HHHHHH). The code for Level Flight is "LVL". Level flight is assumed if allowed and not otherwise specified by a player when he begins to move an aircraft.
}{
In level flight, all FPs must be expended as HFPs. The code for level flight is “LVL.” Level flight is assumed if allowed and not otherwise specified by a player when they begin to move an aircraft.
}

\x{
Example Of Level Flight: An aircraft has a speed of 3.0. The player moves the aircraft through three hexes as shown \changedin{1C}{1C-figures}{below}{in Figure~\ref{figure:level-flight}.}\deletedin{1C}{1C-figures}{\addedin{1B}{1B-apj-23-errata}{\ [The diagram actually shows the left aircraft moving four hexes/hexsides. This is in error.]}} Its turn is over when it has finished moving.
}{
Example: An aircraft has a speed of 3.0. The player selects level flight. The aircraft has three HFPs and moves through three hexes, as shown in Figure~\ref{figure:level-flight}. The aircraft’s move is over after it has expended all of its three FPs.
}

\input{figures/figure-level-flight}

\paragraph{Climbing Flight.} 
\x{
In climbing flight, the aircraft selects a specific type, either a Sustained, Zoom, or Vertical climb, and determines the altitude gain in levels allowed for each VFP expended. A mix of HFPs and VFPs can then be expended subject to the limits outlined in the Climbing rules of section 8.
}{
In climbing flight, the aircraft selects a specific type, either a sustained, zoom, or vertical climb, and determines the altitude gain in levels allowed for each VFP expended. A mixture of HFPs and VFPs can then be expended subject to the limits outlined in rule~\ref{rule:climbing-flight}.
}

\paragraph{Diving Flight.} 
\x{
In diving flight, the aircraft selects a specific type, either an Unloaded, Steep, or Vertical Dive, and determines the altitude loss in levels allowed for each VFP expended. As with Climbing flight, the Diving Flight rules include limits on how many of each type of FP may be expended in a turn depending on the type of dive but some mix of HFPs and VFPs can be expended.
}{
In diving flight, the aircraft selects a specific type, either an unloaded, steep, or vertical Dive, and determines the altitude loss in levels allowed for each VFP expended. A mixture of HFPs and VFPs can then be expended subject to the limits outlined in rule~\ref{rule:diving-flight}.
}

\x{
\paragraph{Expending Mixed FPs.}
HFPs and VFPs may be intermixed and expended in any order so long as any limits for the aircraft's actual climb or dive type chosen are adhered to.
}{
\paragraph{Order of HFPs and VFPs.}
HFPs and VFPs may be expended in any order so long as any limits for the aircraft’s specific climbing or diving flight type are obeyed (although see advanced rule~\ref{rule:changing-flight-type}).
}

\x{
Example Of Altitude Changing Flight: Assume an aircraft with a speed of 7.0 is allowed to expend up to two thirds of its FPs as VFPs. The player could have up to five VFPs and two HFPs but elects only to use three of its FPs as VFPs and the rest as HFPs. He may expend the seven FPs as follows:
HHHHVVV, or VVVHHHH, or HHVVHVH, or in any other order desired.

Note: No matter how the FPs are mixed, the aircraft will only be moved horizontally four hexes as all other FPs will be used to change altitude.
}{
Example: Assume an aircraft with a speed of 7.0 is allowed to expend up to two-thirds of its FPs as VFPs. The aircraft could have five VFPs but elects to use three VFPs and four HFPs. The aircraft may expend the seven FPs as follows: HHHHVVV, VVVHHHH, HHVVHVH, or in any other order. No matter in which order the FPs are used, the aircraft will only move horizontally four hexes as all the other FPs are VFPs that change altitude but do not move the aircraft horizontally.
}

\paragraph{Abnormal Flight.} 
\x{
Two other types of flight are possible. These are Stalled and Departed Flight (see rule 6.4).  An aircraft in abnormal flight may not select Level, Climbing, or Diving flight until recovered from the abnormal condition.
}{
Two other types of flight are possible. These are stalled flight and departed flight (see rule~\ref{rule:abnormal-flight}
) and are considered abnormal flight. An aircraft in abnormal flight may not select level, climbing, or diving flight until it has recovered from the abnormal condition.
}

\begin{advancedrules}

\x{
\section{Half Flight Points}
}{
\section{Half FPs}
}

\notein{1B}{FH in APJ 36 states “If an A/C has a fractional speed and a 0.5 FP carry available, it must use the carry to create a full HFP or VFP.” I don't see what needs changing in the text of the rules.}

\x{
Rather than ignoring left over half flight points, they may take into account as explained below.
}{
Rather than ignoring half FPs, they may take into account, as explained below.
}

\x{
\paragraph{Half FPs.} 
Any 0.5 VFP or 0.5 HFP not spent during a game turn is carried forward to the next game turn as a generic half FP. This is noted in the 0.5 FP Carry line of the Aircraft Log of the coming game turn. The existence of a carried half FP does not change the aircraft's new start speed for the next turn, but, if the new start speed has a half FP in it, it will marry up with any carried half FP to provide another whole FP.
}{
\paragraph{Half Flight Points.} 
Any half FP not expended during a game turn is carried forward to the next game turn. This is noted in the 0.5 FP carry line of the next game turn in the aircraft log sheet. A carried half FP does not change the aircraft's new start speed for the next game turn, but if the new start speed has a half FP, it will marry up with any carried half FP to provide another whole FP.
}

\x{
For example, an aircraft with a start speed of 6.5 having a carried 0.5 FP in the Carry line of the Aircraft Log has 7.0 FPs to expend that game turn. For all applicable game purposes, its speed is still 6.5. If a 0.5 FP carry cannot be mated to a half FP in the start speed, it may be carried forward again (and again) until used. When used, it is gone.
}{
Example: An aircraft with a start speed of 6.5 and a carried half FP from the previous game turn has 7 FPs to expend in the current game turn. For all other game purposes, its speed is still 6.5. If a carried half FP cannot be mated to a half FP in the start speed, it is carried forward again indefinitely until used. When it is used, it is gone.
}

\x{
\section{FP Expenditure Restrictions}
}{
\section{Changing Flight Types}
}
\label{rule:changing-flight-type}

\x{
To reflect the distance traveled forward while an aircraft is raising or lowering its nose, before any altitude change can occur, use the following restrictions.
}{
To reflect the time and distance required to change an aircraft's pitch (that is,  raise or lower its nose), when an aircraft changes flight type, the following restrictions apply.
}

\paragraph{After Level Flight.} 
\x{
If an aircraft chooses Diving or Climbing flight, and in the previous game turn it used Level flight, the first FP used in the current game turn must be an HFP. The remaining FPs may be mixed normally.
}{
If an aircraft chooses diving or climbing flight, and it used level flight in the previous game turn, the first FP used in the current game turn must be an HFP. The remaining FPs may be mixed normally.
}

\paragraph{After Climbing or Diving:} 
\x{
If an aircraft chooses Diving or Climbing flight, and in the previous game turn it used the opposite (i.e., last turn = climb, this turn = dive), then HFPs equal to at least 1/2 the aircraft's speed (round down) must be expended before any VFPs can be used (representing distance flown while reversing nose attitude). Exception: High Pitch Rate capable aircraft need only expend HFPs equal to 1/3 their speed (round down) before using VFPs.
}{
If an aircraft chooses diving flight and in the previous turn used climbing flight or if it chooses climbing flight and in the previous turn used diving flight, then it must expend HFPs equal to at least 1/2 of its speed (round down) before using any VFPs. 

There are two exceptions to this rule:
\begin{itemize}
\item If the ADC notes that an aircraft has a high pitch rate (HPR), it need only expend HFPs equal to at least 1/3 of its speed (round down) before using any VFPs.
\item If an aircraft uses a half-role and dive (see rule~\ref{rule:half-rolls-and-dives}) to enter a vertical dive from a zoom or sustained climb, it must expend all of its HFPs (1/3 of its FPs, according to Table~\ref{table:fractions}) before using any VFPs.
\end{itemize}
}

\paragraph{Same Flight Type.}
\x{
If an aircraft continues a climbing or diving type of flight from one turn to the next, then its HFPs and VFPs may be expended in any order.
}{
If an aircraft continues climbing or diving flight from one turn to the next, then it may expend HFPs and VFPs in any order.
}

\Dx{
\section{Formation Flying}

A formation is in effect when one or more aircraft fly and/or operate together as a unit while in close proximity to each other. Formations usually allow for better teamwork.

\paragraph{Formation Types.} Two types of formations are possible: Close and Tactical. Close formations are designated by stacking aircraft in the same hex. Tactical formations exist whenever certain spacing conditions are met.

\paragraph{Formation Size.} Formations can be of the following sizes:

\begin{itemize}
    \item Section (or Element): Two aircraft; one is the leader and the other is the wingman.
    \item Division (or Flight): Three or four aircraft. One is the leader, and the others are wingmen. Alternately, a division may consist of two sections.
\end{itemize}

\paragraph{Formation Leaders.} Each formation must have a designated leader. Leaders are chosen at the start of play or indicated in the scenarios.  If not defined in the scenarios, one division leader is allowed for each four jets in play, and one section leader for each two jets. A division leader doubles as a section leader. During play, formations may split or form according to the leaders available.

\paragraph{Loss of Formation Leaders.} Wingmen always begin play as part of a particular formation. If their original formation leader is lost and no other qualified leader in that formation exists, that formation is dissolved. The former wingmen may join other formations by moving into formation parameters on those leaders. They will not get the initiative benefits but do avoid any penalty for not being in formation.

\subsection{Close Formations}
\label{rule:close-formations}

Up to four aircraft may stack together in the same hex at the same altitude as a close formation. The close formation stack is moved as a single entity when it is the formation leader's time to move. All aircraft in the slack fly exactly as the leader does and must maintain the same speed, altitude and facing as the leader.

\paragraph{Forming Close Formations.} A close formation may be formed prior to the beginning of a scenario (during set up), or in the Aircraft Admin Phase of a game turn whenever two, three, or four friendly aircraft end up in the same position with the same exact facing, speed and altitude.

\paragraph{Splitting Close Formations.} A close formation may split up by declaring the intent to detach airplanes from the stack when it moves. Aircraft which detach are left in the starting hex while the rest of the formation executes its flight. The detached aircraft are then flown.

For example, a four-plane division wishes to split into two sections. Two airplanes (one being a section leader) detach and are left in place while the original division leader and his remaining wingman move. The two detached aircraft may now move elsewhere as a close formation, or split up individually.

A close formation may contain aircraft of differing types or configurations. When an aircraft in the formation is unable to match the leader's moves or speeds, it must detach itself from the formation. Close formations may place restrictions on the activities of wingmen and on the maneuverability of the leader.

\subsection{Tactical Formations}

A tactical formation exists anytime a designated formation leader and his wingmen meet the following parameters:

\begin{itemize}
    \item When they are within six hexes of each other,
    \item and within three altitude levels of each other,
    \item and the wingmen’s' facings are no more than 60{\deg} different from the leader's facing (left or right),
    \item and the leader is not in the wingman's blind arc.
\end{itemize}

Tactical formations can be formed or broken at any time during play by simply flying the aircraft into or out of these established parameters. A tactical formation cannot have more than four aircraft in it. There are no maneuver or combat restrictions in a tactical formation.
}

\end{advancedrules}