\rulechapter{Air-to-Air Missile Flight and Combat}
\label{rule:air-to-air-missiles}

\x{
This chapter describes the procedures for launching and moving missiles in flight, and for conducting missile attacks when a target aircraft is reached. Aircraft may launch missiles against a properly detected and tracked target. The precise requirements for detection and tracking are specific to each type of missile and are described later.
}{
This rule describes the procedures for launching and flying air-to-air missiles and conducting missile attacks on a target aircraft. Aircraft may launch missiles against an appropriately detected and tracked target. The precise requirements for detection and tracking are specific to each type of missile and are described in rules~\ref{rule:infra-red-heat-seeking-missiles} and \ref{rule:radar-guided-missiles}
}

\x{
\section{Missile Data Tables}
}{
\section{Missile Data}
}
\label{rule:missile-data}

\x{
The information required to launch and fly each type of air-to-air missile is provided in \changedin{1C}{1C-tables}{the Missile Data Table (MDT) included in the play aid charts}{Table~\ref{table:missile-data}}. The data included in the table is as follows:

\begin{itemize}

    \item \itemparagraph{Type.} The name and/or model number of the missile.

    \item \itemparagraph{Year.} The operational service entry date, if known.

    \item \itemparagraph{Weight.} Weight in pounds of missile (for load limits).

    \item \itemparagraph{Load.} The load points of missile (for conf.\ limits).

    \item \itemparagraph{Seeker.} A letter code indicating the type of seeker the missile has. E, I, M, A, are infra-red seekers (Early, Improved, Modern, Advanced).  BR, RH, AH are radar seekers (Beam-Rider, Radar-Homer, Active-Homer). A dash means it is unguided and has no seeker.

    \item \itemparagraph{Launch G.} This is the maximum turn rate an aircraft may use during flight to launch the missile without penalty. If a higher rate was used, a cumulative $+2$ modifier must be applied to the roll for each step of higher turn rate.

    \item \itemparagraph{Launch Roll.} The die roll or less needed to successfully launch missile. \changedin{1C}{1C-tables}{The missile combat charts indicate}{Table~\ref{table:missile-launch-modifiers} indicates} the various modifiers to the roll.

    \item \itemparagraph{Turn Ability.} Missiles use \changedin{1C}{1C-tables}{the Turn Charts}{Table~\ref{table:turn} to determined their turn requirements} just as aircraft do. Their maximum allowed turn rates are listed here.

    \item \itemparagraph{Flight Time.} This is the maximum number of game turns the missile may be in play.

    \item \itemparagraph{Visibility.} This is the missile's visibility number used for sighting purposes.

    \item \itemparagraph{ECCM \#.} The missile's resistance rating to jamming and ground clutter. It is used as a die roll modifier in certain situations explained later, see ECM rules.

    \item \itemparagraph{Chaff \#.} This is the missile's vulnerability rating to chaff and mini-jammer decoys. See ECM rules.

    \item \itemparagraph{Flare \#.} This is the missile's vulnerability rating to flare decoys. See ECM rules.

    \item \itemparagraph{Launch Envelope.} The missile's minimum and maximum ranges in terms of hexes that a missile may be fired at a target are listed here for front, side and rear shots (note, for radar missiles the maximum range is a function of the aircraft's radar tracking strength).

    Note: Front shots are those taken from the target's 150 and 180 arcs; side shots are those from the 120 and 90 arcs; rear shots are those from the 60 or less arcs. NA indicates the shot is not allowed. Each two altitude levels difference in altitude between firer and target equal one hex in range.

    \item \itemparagraph{Speed.} This is the missile's BASE speed which is used as described below to determine a missile's actual speed on its first turn of flight. After the first game turn, the missile's speed will be vary based on numerous factors explained below. A slash with a second number indicates the missile has a sustainer motor which lasts a number of game turns equal to the second number.

    \item \itemparagraph{Active Homing.} This is the range in hexes at which an AH missile's own radar seeker becomes active allowing it to track its target.

    \item \itemparagraph{Home-On-Jam.} A “Y” indicates the missile has a home on jam ability which allows it to guide against aircraft using barrage jamming, see ECM rules.

    \item \itemparagraph{Look Down.} A “Y” indicates the missile has the ability to be used in look down parameters in conjunction with a look down capable radar.

    \item \itemparagraph{Roll To Hit.} The two columns list the number or less from the missile's attack die roll that will result in a direct (first column) or proximity (second column) hit. A direct hit gives a minus 2 damage roll modifier for blast effects.

    \item \itemparagraph{Attack Rating.} The two columns list the attack ratings for a direct (first column) or  proximity (2d column) hit for use with the Aircraft damage tables.

\end{itemize}

Notes: Any special characteristics or missile notes are indicated in the notes section and the note text can be found at the bottom of the MDCs.

}{
Table~\ref{table:missile-data} gives information for carrying, launching, flying, and attacking with air-to-air missiles. The data included in the table is as follows:

\begin{itemize}

    \item \itemparagraph{Year.} The operational service entry date.
    
    \item \itemparagraph{Type and Name.} These give the designation or model of the missile and its name.

    \item \itemparagraph{Weight.} This gives the missile weight in pounds (for load limits).

    \item \itemparagraph{Load.} This gives load points of the missile (for configuration limits).

    \item \itemparagraph{Seeker.} A letter code indicating the type of seeker the missile has: E, I, M, and A are early, improved, modern, and advanced infra-red seekers; BR, RH, and AH are beam-rider, radar-homing, and active-homing seekers; and a dash means that the missile is unguided and has no seeker.

    \item \itemparagraph{Launch Turn Rate.} This is the maximum turn rate an aircraft may use during flight to launch the missile without penalty. The aircraft used a higher rate, a cumulative $+2$ modifier applies to the launch roll for each level of higher turn rate.

    \item \itemparagraph{Launch Roll.} The die roll or less needed to successfully launch a missile. Table~\ref{table:missile-launch-modifiers} gives the various modifiers to the roll.

    \item \itemparagraph{Turn Rate.} Missiles use Table~\ref{table:turn} to determine their turn requirements, just as aircraft do. This gives their maximum turn rate.

    \item \itemparagraph{Flight Time.} This is the maximum number of game turns the missile may be in play.

    \item \itemparagraph{Visibility.} This is the missile's visibility number for sighting purposes.

    \item \itemparagraph{ECCM Rating.} This is the missile's resistance rating to jamming and ground clutter (see rule~\ref{rule:electronic-warfare} and \ref{rule:radar-ground-clutter}).

    \item \itemparagraph{Chaff Rating.} This is the missile's vulnerability rating to chaff and mini-jammer decoys (see rule~\ref{rule:electronic-warfare}).

    \item \itemparagraph{Flare Rating.} This is the missile's vulnerability rating to flare decoys (see rule~\ref{rule:electronic-warfare}).

    \item \itemparagraph{Launch Envelopes.} These are the minimum and maximum ranges at which a missile may be launched at a target. (For BRMs and RHMs, the maximum range also depends on the aircraft's radar tracking strength.)

    Front shots are those taken from the target's 150{\deg} and 180{\deg} arcs; side shots are those from the 120{\deg} and 90{\deg} arcs; rear shots are those from the 60{\deg} or 30{\deg} arcs. A dash indicates the shot is not allowed.

    \item \itemparagraph{Speed.} This is the missile's base speed, which is used as described below to determine a missile's actual speed on its first turn of flight. After the first game turn, the missile's speed will vary based on numerous factors explained below. A dash with a second number indicates the missile has a sustainer motor, which lasts a number of game turns equal to the second number.

    \item \itemparagraph{Active Homing.} This is the range in hexes at which an AH missile's radar seeker becomes active, allowing it to track its target.

    \item \itemparagraph{Hit Roll.} The two columns list the number or less for the missile's attack hit roll that will result in a direct (first column) or proximity (second column) hit. A direct hit gives a $-2$ damage roll modifier.

    \item \itemparagraph{Attack Rating.} The two columns list the attack ratings for a direct (first column) or proximity (second column) hit for use with the aircraft damage tables.

    \item \itemparagraph{Notes.} The notes indicate whether the missile has special characteristics such as:
    \begin{itemize}
        \item NW: nuclear warhead (see advanced rule~\ref{rule:air-to-air-nuclear-rockets}).
    \item UNC: compatible with IR uncage technology (see rule~\ref{rule:ir-uncage}).
    \item HOJ: home-on-jam seeker (see rule~\ref{rule:home-on-jam-seeker}).
    \item IA: instant-arming missile (see rule~\ref{rule:instant-arming-missiles}).
    \item LD: look-down missile (see rule~\ref{rule:look-down-missiles}).
    
    \end{itemize}

\end{itemize}
}

\section{Missile Launches}
\label{rule:missile-launches}

\x{
An aircraft may launch up to two missiles at a targeted enemy aircraft during the Air to Air Missile Launch Phase.
}{
An aircraft may launch up to two missiles at an enemy aircraft during the air-to-air missile launch phase.
}

\addedin{1C}{1C-tables}{
    \input{tables/table-missile-launch-modifiers}
}

\Ax{
\paragraph{Missile Launch Requirements.} 
The following requirements apply to all types of missile:
\begin{itemize}

    \item All missiles launched by an aircraft in a single phase must be of the same type (for example, AIM-9 or AA-7). An aircraft may launch IR and RH versions of a single type in the same phase if the targeting requirements of each seeker type are met.
    \item An aircraft that engaged a missile (see rule \ref{rule:engaging-missiles}) may not launch missiles.
    \item An aircraft that turned at the ET rate may not launch a missile until after a recovery period (see rule~\ref{rule:recovery-periods}).

    \item An aircraft that fired its guns or rockets at after its last FP that game turn may not launch missiles.

    \item An aircraft that prepared for or executed a roll maneuver during its last FP that game turn may not launch missiles.
    \end{itemize}

The following requirements apply to IR missiles:
 
\begin{itemize}

    \item If an aircraft used climbing flight during the game turn of launch, it may not fire at a lower target.

    \item If an aircraft used diving flight during the game turn of launch, it may not fire at a higher target .

    \item If the launching aircraft used level flight during the game turn of launch, the difference in altitude between the firer and target may not be more than one level for each two full hexes of horizontal range.

\end{itemize}

The following requirements apply to BR, RH, and AH missiles:

\begin{itemize}

    \item An aircraft may not fire on a the target that is not within its radar arc limits.

\end{itemize}
}

\x{
\paragraph{Procedure.} 
Announce the number of missiles that each aircraft will attempt to fire (one or two) and their intended target during the Air to Air Missile Launch Phase.

Roll the die for each missile.\addedin{1C}{1C-tables}{\ Table~\ref{table:missile-launch-modifiers} shows the modifiers to the launch roll.} If the result is equal to or less than the missile launch roll number given for that type of missile \changedin{1C}{1C-tables}{on the MDT}{in Table~\ref{table:missile-data}}, the launch is successful, place a missile counter in the launch aircraft's position with the exact same facing and altitude as the aircraft. On the next game turn, the missile will fly in pursuit of its target. \changedin{2A}{2A-missile-launch}{If the launch roll is greater than the launch number}{If the launch roll fails by more than one or is an unmodified 10}, the missile has malfunctioned (through faulty guidance, or because of a dud motor). The malfunctioned missile is removed from play. In either case, the missile and its load points are considered expended. \addedin{2A}{2A-missile-launches}{If the launch roll fails by exactly one and is not an unmodified 10, the launch fails, but the missile remains on the rail and can be used in the future.}

If all declared launch attempts from an aircraft fail, the player may attempt one additional launch that phase. If at least one missile fired previously, this last attempt is not allowed. A missile's target is noted when it is launched.
}{
\paragraph{Missile Launch Procedure.} 
Announce the type and number of missiles each aircraft will attempt to fire (one or two) and their target.

Roll the die for each missile.\addedin{1C}{1C-tables}{\ Table~\ref{table:missile-launch-modifiers} shows the modifiers to the launch roll.} 

\begin{itemize}
    \item If the modified result is equal to or less than the missile launch roll number given in Table~\ref{table:missile-data} \emph{and is not an unmodified 10}, the launch succeeds. Place a missile counter at the launch aircraft's position with the same facing and altitude as the aircraft. On the next game turn, the missile will fly in pursuit of its target. 

    \item If the modified result is only one greater than the missile launch roll number given in Table~\ref{table:missile-data} \emph{and is not an unmodified 10}, the launch fails.
    In this case, the missile is not launched but remains on the rail and can be used in a subsequent game turn.

    \item If the modified result is more than one greater than the missile launch roll number given in Table~\ref{table:missile-data} \emph{or is an unmodified 10}, the missile fails (perhaps because of faulty guidance or a dud motor). The failed missile is removed from play.  


\end{itemize}

In the first and third cases, the missile is considered to have been expended, and its load points are no longer counted for the aircraft's load.

If all declared launch attempts result in launch failures or missile failures, the player may attempt one additional launch in that phase. This last attempt is not allowed if at least one launch attempt was successful. 

A missile's target is noted when it is launched.
}
\Dx{
\paragraph{Missile Launch Restrictions.} Only “free” aircraft may launch missiles. All missiles launched by one aircraft in a single phase must be of the same type (for example, AIM-9 or AA-7). IR and RH versions of a single type may be launched in the same phase by the one aircraft if the targeting requirements of each seeker type are met.

Aircraft attempting to launch Visually aimed (IR) missiles are restricted as follows:
 
\begin{itemize}

    \item If the launch aircraft climbed during the turn of launch, it may not fire at a lower target.

    \item If the launching aircraft dove, it may not fire at a higher target.

    \item If the launching aircraft flew level, the difference in altitude between the firer and target may not be more than one level for each two hexes distance between the two.

\end{itemize}

Aircraft attempting to launch Radar Guided missiles are restricted as follows:

\begin{itemize}

    \item The target must be within the aircraft's radar arc limits or the launch aircraft may not fire.

\end{itemize}

Aircraft attempting to launch any type of missile are restricted as follows:

\begin{itemize}

    \item An aircraft that turned at ET may not launch a missile unless the recovery period described in chapter 9 has been met.

    \item An aircraft that fired its \changedin{2B}{2B-launch-restrictions}{guns during}{guns or rockets after} its last FP that turn may not launch missiles.

    \item An aircraft that executed a roll maneuver, or prepped for a roll during its last FP that turn, may not launch missiles.
    
\end{itemize}
}

\paragraph{Missile Launch Modifiers.} 
\x{
High aircraft turn rates during the turn of launch can adversely affect a missile's ability to stay locked on to a target during its separation from the launch aircraft and motor ignition. Each missile has a Launch G (turn rate) listed for it \changedin{1C}{1C-tables}{on the MDT}{in Table~\ref{table:missile-data}}. This is the highest turn rate the launching aircraft may use in the game turn and still be able to launch those missiles without penalty. If a higher turn rate was used in the turn, a $+2$ modifier is applied to the launch roll for each turn rate above the listed Launch G.

For example, if a missile has a Launch G of “TT”, and the aircraft used a BT turn rate during the game turn, a $+4$ modifier to the launch roll would apply.

Additional launch roll modifiers may exist for weather, terrain clutter, crew quality, and the presence of missile countermeasures in the form of decoys (expendable chaff, flares, and mini-jammers), or electronic jamming. These modifiers are explained in later rules.
}{
Table~\ref{table:missile-launch-modifiers} gives modifiers for  missile launch rolls:
\begin{itemize}
    \item Aircraft turn rates during launch can adversely affect a missile's ability to stay locked-on to a target during its separation from the aircraft and motor ignition. 
    
    Each missile has a maximum launch turn rate given in Table~\ref{table:missile-data}. If the launching aircraft has used a higher turn rate and has not yet completed the corresponding recovery period (see rule~\ref{rule:recovery-periods}), it suffers a $+2$ modifier for each turn rate above the maximum.

    For example, if a missile has a maximum launch turn rate of TT and the launching aircraft used a BT turn rate and has not yet completed the recovery period, a $+4$ modifier would apply to the launch roll.

    \item If a missile is launched out-of-envelope, it receives a $+3$ modifier (see advanced rules \ref{rule:irm-out-of-envelope-launches} and \ref{rule:radar-out-of-envelope-launches}).
    
    \item If an IRM is at a target in ground clutter, it suffers a $+2$ modifier (see rule \ref{rule:irm-ground-clutter}).

    \item If an IRM is launched above the highest cloud layer at a lower target during daylight, it suffers a $+3$ launch modifier (see advanced rule \ref{rule:clouds}).

    \item If an IRM is launched into Sun clutter, it suffers a $+3$ launch modifier (see advanced rule~\ref{rule:sun}).

    \item If an IRM is launched at a target with an active DDS program, it suffers a launch modifier equal to the smaller of the DDS flare PPL and the missile flare vulnerability rating (see advanced rule~\ref{rule:dds}).

    \item If a BRM, RHM, or ARM is snap-fired, it suffers a $+2$ launch modifier (see advanced rule~\ref{rule:snap-firing-missiles}).
    
    \item The quality and characteristics of the crew give launch modifiers (see advanced rule~\ref{rule:crew-quality} and \ref{rule:crew-characteristics}). 
    
    In multi-crew aircraft, the pilot is the relevant crew member for IRM launches and the radar operator (RIO, WSO, or equivalent) is the relevant crew member for BRM, RHM, and ARM launches.

    \item If the launch aircraft is damaged, it suffers a launch modifier.
    
\end{itemize}
}
\section{Missile Flight}
\label{rule:missile-flight}

Missile flight has been simplified for ease of play. A missile must expend all of its FPs each turn. It does not differentiate between HFPs and VFPs as aircraft do. The missile's start speed and altitude is recorded on the aircraft log each game turn. Unlike aircraft, missiles may both climb and dive (within restrictions) in the same turn in order to pursue their target. Missiles never have or carry half FPs or partial altitude gains.

\paragraph{Start Speeds And Altitudes}. On its first turn of flight (the game turn after the one in which it was launched) a missile's start altitude is the same as the launch aircraft's start altitude the turn after launch. Its start speed is its base speed listed \changedin{1C}{1C-tables}{on the MDT}{in Table~\ref{table:missile-data}} adjusted as follows:

\begin{itemize}

    \item minus one if the speed of the launch aircraft on the turn after launch is 3.0 or less or,

    \item plus one if the start speed is 6.0 or more or

    \item plus two if the start speed is 9.0 or more
\end{itemize}

On subsequent turns, the missile's start altitude will be the altitude it ended up at after its flight in the previous turn and its start speed will be two thirds its previous speed rounding fractions up, and adjusted as follows:

\begin{itemize}

    \item minus one if the missile's altitude gain for the turn equaled half or more of its speed in levels, or minus two if it equaled its speed or more in levels.

    \item plus one if the missile's altitude loss for the turn equaled half or more of its speed in levels, or plus two if it equaled its speed or more in levels.

\end{itemize}

Example: A missile with a start speed of 14 gained 8 levels during its flight. Its new start speed would be 8 (2/3d's of 14 = 9; minus 1 for altitude gain).

\paragraph{Sustainer Motors.} Missiles equipped with sustainer motors (indicated \changedin{1C}{1C-tables}{on the MDT}{Table~\ref{table:missile-data}}) determine their start speed differently. If a missile has a sustainer, its first turn of flight speed is equal to its base speed as indicated \changedin{1C}{1C-tables}{on the MDT}{in Table~\ref{table:missile-data}} plus the full speed of the aircraft (rounding fractions up). For every turn the sustainer lasts, the missile's speed is not reduced to two thirds as above but only adjusted for climbs and dives. After the sustainer burns out, the missile's speed is reduced as above. A missile with a sustainer will have a number after a slash in the Speed column \changedin{1C}{1C-tables}{of the MDT}{in Table~\ref{table:missile-data}}. The number indicates how many turns the sustainer motor lasts.  A “1” indicates the sustainer is only good for the first turn of flight.

\paragraph{Missile Order of Flight.} A missile always flies at the same time as its target.

\paragraph{Missile Proportional Flight.} Missiles and their targets move simultaneously. To simulate this, the missile and its target always alternate their expenditure of FPs in proportion to their relative speeds\addedin{2A}{2A-missile-flight}{, with the missile moving first}. Proportional speed is determined by dividing the faster speed by the slower speed. The result (ignoring fractions) is the number of FPs expended by the faster missile per FP expended by the slower aircraft.  Leftover missile FPs are tacked on as extra FPs in the first few segments of the proportional flight.

For example, if the aircraft speed is 5.0 and the missile speed is 15.0. Five goes into fifteen three times. Thus, the missile will move 3 FP for each 1 FP moved by the aircraft.  In this case, \changedin{2A}{2A-missile-flight}{the aircraft would move one FP and then the missile would move three}{the missile would move three FPs and then the aircraft would move one}.  This continues until both are out of FPs for the turn or the missile reaches a position from which an attack can occur.

If the proportion includes leftover FPs, for example, say the missile's speed was 17.0 instead of 15.0. In this case five goes into seventeen three times with two left over. The missile would still move three for one except the two extra points are added on one at a time in the first two segments of movement that turn. That is, \changedin{2A}{2A-missile-flight}{the aircraft would move one, then the missile four ($3+1$), then the aircraft one and missile four again}{the missile would move four FPs, then the aircraft one, then the missile four, and the the aircraft one again}. After that, the missile reverts to its three to one proportion.

If an aircraft's speed includes a fraction; say it is 4.5 instead of 5.0, round it up to simplify determining proportions but use its actual speed when moving. This would give the same result as in the above paragraph (5 into 17). The difference comes in the execution of the segments. The moves would be 4 for 1, 4 for 1, 3 for 1, 3 for 1, and then 3 for the aircraft's half FP.

\deletedin{2A}{2A-spbr}{
\paragraph{Speedbrake Effects on Proportional Moves (adv. rule 6.5)} When engaged in proportional movement with attacking missiles, speedbrakes (when applied) can only use up the last available FP to be expended. Or, stated another way, when engaged in proportional flight with attacking missiles, if the aircraft still has at least one full FP to expend it must do so each on every proportional move whether speedbrakes have been used yet or not.
}

\paragraph{Missile Types Of Flight.} A missile may fly level, climb and dive all in the same game turn as necessary to pursue its target. The following \changedin{1C}{1C-tables}{is}{and Table~\ref{table:missile-flight-rules} are} a summary of missile types of flight.

\addedin{1C}{1C-tables}{
    \input{tables/table-missile-flight-rules}
}

\begin{itemize}

    \item \itemparagraph{Level Flight.} The missile expends one FP to fly forward one hex or hexside. A missile may freely lose one altitude level (at no FP cost) for each hex/hexside it enters.

    \item \itemparagraph{Climbing Flight.} The missile expends one FP to climb one or two levels. A missile may not climb if its target is at the same or a lower altitude level. Exception: Those missiles noted as “TVM” (track-via-missile) or “MCG” (Mid-Course-Guidance) capable may climb as high as desired above a target before diving back down to intercept. The maximum altitude level any missile in play is allowed to reach is 100.

    \item \itemparagraph{Diving Flight.} The missile expends one FP to dive two or three levels. A missile may not dive if its target is at the same altitude level or a higher one. Once per game-turn, a missile may lose just one altitude with one FP. Remember, missiles may lose altitude freely by one level, for each hex entered in level flight mode.

    \item \itemparagraph{Combined Flight.} FPs expended to move forward or to change altitude may be intermixed in any order. Any number of FPs may be expended climbing OR diving on the same hex/hexside. However, a missile may not both climb AND dive from the same hex/hexside and the following restrictions apply if the missile wishes to switch directly from climbing to diving or vice versa:

    \begin{itemize}

        \item \changedin{1B}{1B-apj-23-errata}{if the missile’s Turn Ability is less than BT/2, it must expend two FPs in forward level flight (free altitude loss not allowed for these) between climbing and diving FPs.}{If the missile's Turn Ability is not BT/2, ET/2, or ET/3 it must expend FPs in level flight equal to {\onethird} its speed (round up) between VFPs used to climb and dive (free altitude loss not allow for these either).}
        
        \item \changedin{1B}{1B-apj-23-errata}{if the missile’s Turn Ability is BT/2 or greater (e.g., ET/2 or ET/3), only one FP must be expended in level flight between the climbing and diving FPs.}{If the missile's Turn Ability is BT/2 or better, it need only FPs equal to $1/10$ its speed (round up) in level flight between VFPs used to climb and dive.}

    \end{itemize}

    \item \itemparagraph{Turning.} Missiles use the Turn Charts just as do aircraft. Missiles do not receive decel points for turning or changing facing. Missiles never consider angles of bank and may therefore reverse turns instantly.\deletedin{2A}{2A-snap}{ Missiles may use a snap turn once in their entire flight. If the snap turn is the first action performed, the normal prep-move requirements for snap-turning are waived (see Arming below).} If a divisor is given, then the turn rate is better by the divisor's factor.

    Example: A missile with ET/2 turn rate requires only half the normal ET requirement. At speed 16.0, in the LO band, it would only have to move 3 FPs per facing change (half of six). Always round fractions up when determining missile turn requirements (i.e., half of 5 would be 3, or a third of 7 would be 3).

    \item \itemparagraph{Maneuvers.} Missiles are allowed slide maneuvers on every game turn like aircraft, lag displacement, and barrel rolls are not allowed. Any missile which expends more than one FP from the same position while climbing or diving may execute a vertical roll. Missiles are limited to one vertical roll in their entire flight unless they are pursuing a target which also does vertical rolls. In this case, the missile is allowed as many additional vertical rolls in a game turn as their target has performed.

    The normal prep-move requirements for all maneuvers must be met just as for aircraft and all maneuvers cost 1 FP to execute except for vertical rolls which cost 0. \addedin{1B}{1B-apj-23-errata}{Missile\deletedin{2A}{2A-snap}{ snap turns (except those performed as Instant Snap Turns) and missile} slides are subject to a $+1$ prep requirement if the missiles are supersonic.}

\end{itemize}

\paragraph{Missile Arming.} Missiles must complete arming before they can turn or maneuver in pursuit of a target. Arming occurs automatically after the missile has flown a certain distance, usually one hex.

To become armed, missiles must expend their first FP moving forward in level flight unless on the turn of launch the firing aircraft was in a climb or dive where more than half the aircraft's flight was spent as VFPs. In this case, the missile could use an FP to gain or lose altitude as appropriate instead. This one FP in which missiles are becoming armed may not be counted toward prep-moves, or turning requirements, nor can the missile attack targets it reaches (they are missed). They may commence maneuvering normally on second and subsequent FPs.                                                          
\paragraph{Missile Tracking.} At the end of each game turn, and at the end of each segment of proportional movement, a missile must have the target within the tracking parameters of its seeker head (in terms of angle-off) or it will lose contact and either self-destruct or enter ballistic unguided flight. In either case, it is removed from play. The exact tracking requirements are described for each missile kind of missile later in the rules.

\changedin{2A}{2A-snap}{
\paragraph{Instant Snap Turning (adv. rule 7.3).} Instant arming missiles, (indicated by an asterisk \changedin{1C}{1C-tables}{on the MDT}{in Table~\ref{table:missile-data}}) may begin maneuvering and attack immediately after launch. They are also allowed to immediately snap turn without prep-moving with their first FP or any FP later in their flight so long as no turns, maneuvers, or switches between climbs and dives have yet commenced. If they elect to snap-turn later in their flight, after any turns, maneuvers, or switches have occurred, they must prep-move for the snap-turn normally.

Non-instant arming missiles may also snap-turn as their first maneuvering actions, but only after becoming armed. This is allowed as above, with their first FP following becoming armed, or later as long as no turns, maneuvers, or switches between climbs and dives have yet occurred or commenced. As above, if they elect to snap turn later in their flight they must prep-turn normally.
}{
\label{rule:instant-arming-missiles}
Instant arming missiles, (indicated by an asterisk \changedin{1C}{1C-tables}{on the MDT}{in Table~\ref{table:missile-data}}) may begin maneuvering and attack immediately after launch.
}

\paragraph{Follow On Missiles.} Whenever two missiles are launched at a time, the second one is termed a “follow-on” missile. This missile may not begin moving until the first missile has moved at least two FPs. This simulates the time delay between the firing of the two missiles. The time delay may be longer, but not more than 1/3 the first missile's speed.

When the follow-on missile first begins moving, it is not allowed to move more FPs than the lead missile has left to it in that segment of the proportional move. For example, if the proportions are 5 to 1 and the follow on starts moving after the lead missile moves three points, the follow on would only be able to move 2 points (the same as was left to the lead). In the next segment, both could move 5 hexes but the follow-on missile would still trail by a distance of 3 (the original delay).

On the first and subsequent turns of flight, the follow-on must cease moving when the first missile does. Thus, on the first turn, it will be cheated out of FPs equal to the delay between it and the first missile. To compensate, the follow-on missile is allowed to use those FPs on the last turn of its flight, they are added as bonus FPs one per proportional segment starting with the first until used up.

\section{Missile Attacks}

\addedin{2A}{2A-missile-attacks}{
    \input{figures/figure-missile-attacks}
}

\notein{1B}{AWF: FH in the APJ 36 errata has text about a missile attacking if the target enters the missile's position, but this is included in the TSOH errata and incorporated in 1B.}

\paragraph{When Does A Missile Attack?} \changedin{1B}{1B-apj-23-errata and 1B-apj-26-qa}{At the instant a missile starts a proportional move with its target in its 180+ arc and the range (in hexes; 2 levels of altitude equals one hex) is equal to or less than the number of FPs the missile has available in that proportional move, a missile attack is declared.}{\changedin{2A}{2A-missile-attacks}{At the beginning of or during a missile's proportional move, the instant the missile has its target in its 180+ arc and the range (in hexes; 2 altitude levels = 1 hex) is equal to or less than the number of FPs the missile has remaining in that proportional move, a missile attack \emph{must} be declared.}{A missile attacks when it passes through the target hex at its altitude or when is co-altitude with the target, has at least one FP remaining in the proportional move, and the target occupies one of the positions show in Figure~\ref{figure:missile-attacks}.} \addedin{1B}{1B-apj-36-errata}{If the missile finishes arming in the target's location, it attacks.} Should a target aircraft enter the missile's position during its move, it is also immediately attacked.} 


\paragraph{Procedure.} Roll the die and apply any modifiers. Compare the modified result to the missile's to hit numbers to determine if a hit was achieved, and if so, what kind of hit. If a hit is achieved, roll on \changedin{1C}{1C-tables}{the Damage Tables}{Table~\ref{table:aircraft-damage}} using the appropriate attack rating.

\addedin{1C}{1C-tables}{
    \input{tables/table-missile-attack-modifiers}
}

The two “Roll to Hit” columns \changedin{1C}{1C-tables}{on the MDT}{in Table~\ref{table:missile-data}} indicate the Die Roll or less result that will give a direct or proximity hit. A direct hit occurs if the roll was equal to or less than the indicated number. A proximity hit (which has a lesser attack rating) occurs if the roll was greater than that for a direct hit but less than or equal to that given for a proximity hit. If the roll was higher than the proximity number, the missile misses.

\paragraph{Roll To Hit Modifiers.} The Missile Attack Modifiers Table lists the modifiers for missile angle off, terrain clutter effects, target considerations, and ECM. For purposes of the angle-off to hit modifiers, the missile is always considered to be in the angle-off arc it was in during the proportional move in which the missile attack was declared. If it is on a line between two arcs, it is considered to be in the arc which favors the defender.

\paragraph{Missile Damage.} The two Attack Rating columns \changedin{1C}{1C-tables}{of the MDT}{in Table~\ref{table:missile-data}} list the attack rating of the missiles for direct and proximity hits respectively. The ratings are used to determine aircraft damage as described in chapter 10. However, a direct hit always gets a minus 2 modifier to the damage roll (warhead blast effects as opposed to just shrapnel from a proximity hit).

\x{
\section{Defensively Engaging Missiles}
}{
\section{Engaging Missiles}
}
\label{rule:engaging-missiles}

An aircraft under attack by one or more missiles may declare itself defensively engaged against any of the missiles which are sighted or to which it has been alerted to by ECM. This declaration is made during the aircraft decisions phase.

\paragraph{Engaged Aircraft Versus Missiles.} An “engaged” aircraft is considered to be actively defending itself against attacking missiles. As such, it is allowed the following benefits during its move:

\begin{itemize}

    \item Idle power may be selected and is automatically effective providing modifiers to IR missile attacks.

    \item Missile decoys (chaff/flare/jammers) may be manually deployed to defeat the missile. Manual decoys may be deployed in addition to any being deployed by a DDS program (see ECM rules).

    \itemdeletedin{2A}{2A-missile-flight}{The target aircraft begins proportional movement first, expending its first FP before the missile expends any FPs.}
    
\end{itemize}

\paragraph{Restrictions.} Engaged aircraft are not allowed to make attacks of any sort and may not launch weapons, use “T”-level flight or do damage control while engaged.

\paragraph{Free Aircraft Versus Missiles.} A free aircraft under attack by a missile has either not spotted it, or opted to ignore it, possibly depending on a dispenser program to stop it. As such it has the following disadvantages:

\begin{itemize}

    \itemdeletedin{2A}{2A-missile-flight}{The missile begins proportional movement first expending its FPs before the aircraft does.}

    \item Free aircraft may not manually deploy decoys, although they may gain the benefit of any DDS programs already in operation.

    \item Idle power selected by a free aircraft is only effective against the missile on a die roll of 1 to 4.

    \item The missile is given an additional $-1$ modifier on the to hit roll.

\end{itemize}

Note: Free aircraft are not restricted like engaged aircraft.

\paragraph{Multiple Missile Attacks.} An aircraft that engages one missile, is considered engaged against all missiles currently pursuing it. However, manual decoys may only be employed against missiles which are sighted or to which it is specifically alerted by ECM. Against engaged aircraft, missiles which were not sighted or alerted to are still affected by Idle power and do not get the $-1$ to hit modifier\deletedin{2A}{2A-missile-flight}{ but still move first as if the target were a free aircraft}.

\section{Missile Countermeasures}

Aircraft defend against missiles by trying to outmaneuver them (not likely) and by using expendable decoys such as chaff, flares and mini-jammers, and through electronic warfare. A “decoy” in the game, represents a cluster of two to four actual expendables.

\paragraph{Out-Maneuvering Missiles.} If an aircraft reaches a position where a missile cannot move to keep it within its tracking requirements at the end of the turn or proportional movement segment, or if the missile is below maneuver speed and the target is not directly in front of it when an attack is declared, the missile is out-maneuvered and removed from play.

\paragraph{Manual Decoy Dispensing.} Aircraft equipped with decoy dispensing systems (DDS) as part of their ECM suite, or aircraft carrying decoy dispenser pods may manually dispense decoys against an attacking missile if they are defensively engaged against it. Manual decoys may be dispensed even if an automatic decoy program is in effect (see rule 19).

When a sighted missile attacks, the engaged aircraft may immediately expend 1 or 2 decoy clusters of each type of decoy available. Quantities dispensed of each type must be equal (although some types may run out early).

The defender rolls the die once for each decoy dispensed. For Chaff or Flares, if the roll is less than or equal to the missile's appropriate decoy vulnerability number (given \changedin{1C}{1C-tables}{on the MDT}{in Table~\ref{table:missile-data}}), it is decoyed and removed from play. For mini jammers, if the roll is less than or equal to the missile's Chaff Vulnerability plus 1, it is decoyed and removed from play. If the missile is not decoyed, it rolls for its attack.

\paragraph{Automatic Decoy Dispensing.} Decoy dispensers may be used to continuously emit decoys over the course of a game-turn via automatic programs as described in rule 19. A decoy program provides a “Protection Level” number which is used as an attack die roll modifier if it is dispensing decoys to which a missile is vulnerable.

\paragraph{Electronic Warfare.} Radar guided missiles and aircraft radars may be vulnerable to electronic jamming which affects their ability to launch and track, and their attack die rolls. Rule 19 covers electronic warfare in detail.

\begin{advancedrules}

\section{Realistic Missile Speed Attenuation}

\addedin{1C}{1C-tables}{
    \input{tables/table-missile-speed-attenuation-factor}
}

Missiles use powerful boost motors which accelerate them to top speed within a matter of two to three seconds. After that, most missiles simply glide to their targets rapidly losing speed along the way.  Some have sustainer motors which burn for a short period of time after the booster goes out and these lose speed at a lesser rate. Nevertheless, the speed loss can be dramatic, up to a third of the missile's top speed within the span of a single turn depending on its actions. This rule replaces the generic method for determining missile speeds given in 14.3.

\paragraph{Missile Speed Attenuation Factor.} To realistically account for the high-speed loss that can occur each turn, a Speed Attenuation Factor is applied to the missile's base start speed to get an average speed for the turn. The average speed indicates the number of FPs that the missile has. The average speed is therefore the speed that is listed on the aircraft log and used for the missile's flight.

\paragraph{Missile Base Start Speed.} On the first turn of the missile's flight, the base start speed equals the missile's listed speed \changedin{1C}{1C-tables}{from the MDT}{in Table~\ref{table:missile-data}} plus the speed of the launching aircraft. On all subsequent game turns, the base start speed is the missile's previous average speed plus any changes for climbing, diving, and maneuvering.

\paragraph{Procedure.} At the beginning of every turn of its flight, including its first, determine a missile's average speed as follows:

\begin{itemize}

    \item Refer to \changedin{1C}{1C-tables}{the Missile Speed Attenuation Table}{Table~\ref{table:missile-speed-attenuation-factor}} and find the altitude band the missile is starting in.

    \item Cross index the band with the game turn of the missile's flight to find the attenuation factor.

    \item Multiply the attenuation factor by the missile's base start speed. Round resulting fractions up at .5 or better, and drop fractions of less than .5.

\end{itemize}

The final result is the missile's average speed for the current turn. Note: \changedin{1C}{1C-tables}{A math saver table is provided in the play aids, which}{Table~\ref{table:missile-speed-math-saver}} does the math for you.

\paragraph{Climbing and Diving Effects On Speed.} At the end of a game turn increase or decrease the missile's speed as given in 14.3 for climbs and dives.

\paragraph{Maneuvering Effects On Speed.} At the end of game turn, reduce the missile's speed by one for each 30 degrees of facing change it accomplished by turning during that turn.

\paragraph{Sustainer Motor Effects.} Some missiles have sustainer motors which provide extra thrust after the missile's booster gives out. This rule replaces the one given in 14.3.

On the first game turn sustainer powered flight, use the missile speed attenuation factor that applies for the altitude band two above that the missile is actually in or the UH band, whichever occurs first, to determine average speed. For each game turn of sustainer powered flight after the first, the speed attenuation factor is 1.0 regardless of its current altitude (meaning no attenuation speed loss applies, though other speed loss effects do). Once the sustainer gives out, normal speed attenuation factors are used.

\paragraph{Minimum, Maneuver, and Maximum Missile Speeds.} \changedin{1C}{1C-tables}{Listed Next to the Missile Speed Attenuation Table is}{Table~\ref{table:missile-speed-limits} gives} the minimum, maneuver and maximum speeds allowed to any missile in a given altitude band. Speed gain above the maximum speed is not allowed, excess speed gain is lost. Any missile with a start speed of less than the minimum listed is considered to stall out and is removed from play. Any missile with a start speed of less than maneuver speed may not turn or perform maneuvers of any sort. It may only fly forward and climb or dive.

\section{Formation Effects on Missile Attacks}

\paragraph{Missiles Versus Close Formations.} Heat seeking missiles launched at a close formation, randomly determine which aircraft is the actual target. Radar, laser, or optically guided missiles target aircraft normally (See following missile rules).

\paragraph{Engaging Missiles.} Aircraft that remain in close formation may not engage missiles. If any aircraft wishes to detach to engage a missile, the close formation is automatically nullified and all must move during the engaged aircraft movement phase and all are restricted from performing actions as if each had engaged the missile (they would all initially be unsure of who the real target is).

\end{advancedrules}
